---
title: "MKM_2749577_RP_Full_Analysis"
author: "Micah Mallée"
date: "2024-06-07"
output: html_document
---

### Load packages
```{r}
library(limma)
library(dplyr)
library(sva)
library(EnhancedVolcano)
library(patchwork)
library(ggpubr)
library(reshape2)
```

### Read annotation and sample information
```{r}
# Read Annotation for all 39428 probes
Annotation <- read.delim("~/SAGA_annotation/Annotation_SAGA_FINAL_20231212.txt", header = TRUE, sep = "\t", stringsAsFactors = FALSE, row.names = 1)

# Read Annotation all 38387 annotated probes
Annotation.known <- read.delim("~/SAGA_annotation/Annotation_SAGA_FINAL_KNOWN_20231212.txt", header = TRUE, sep = "\t", stringsAsFactors = FALSE, row.names = 1)

# Read sample information for known unpublished samples
pData_412 <- read.csv('/home/mmallee/pData_train_412.csv')
pData_258 <- read.csv('/home/mmallee/pData_258.csv')

# Set colors for plotting
condition_colors <- c("A2_Mock" = "#737373", "A1_Mutagenic" = "#CB181D", "A3_Safe" = "#b2df8a", 'A4_Unknown' = '#623456')

class_colors <- c("transforming" = "#CB181D", "untransforming" = "#b2df8a")

ivim_colors <- c("Immortalizing" = "#CB181D", "Safe" = "#b2df8a")

clone_colors <- c("Growth" = "#CB181D", "No_Growth" = "#b2df8a")
```

### Define functions
```{r}
# Creating design matrices per batch
create_design <- function(pdata_obj) {
  groups <- factor(pdata_obj$Design)
  design <- model.matrix(~ 0 + groups)
  rownames(design) <- pdata_obj$`Sample Number`
  if (length(colnames(design)) == 3) {
    colnames(design) <- c("transforming", "mock", "safe")
  } else if (length(colnames(design)) == 1) {
    colnames(design) <- c("transforming")
  } else {
    colnames(design) <- c("transforming", "safe")
  }
  return(design)
}
# Creating contrast matrices per batch

create_contrasts <- function(designs_obj) {
  if (length(designs_obj) == 3) {
    contrast <- makeContrasts(
      c1 = transforming - mock,
      c2 = safe - mock,
      c3 = transforming - (mock + safe) / 2,
      c4 = transforming - safe,
      levels = designs_obj
    )
  } else {
    contrast <- makeContrasts(
      c4 = transforming - safe,
      levels = designs_obj
    )
  }
  return(contrast)
}

GSA.read.gmt <- function(filename) {
  #
  ## Read in and parse a gmt file (gene set file) from the  Broad institute
  # this is tricky, because each lines (geneset) has a variable length
  #  I read the file twice, first to pick up the geneset name and description
  # in the   first two  columns, then I read it all in as a long string

  # The beginning and end of each gene set in the string
  # is determined by matching
  # BOTH on  geneset name and description (since geneset names sometimes
  # occur as genenames elsewhere in the file)

  a <- scan(filename, what = list("", ""), sep = "\t", quote = NULL, fill = T, flush = T, multi.line = F)
  geneset.names <- a[1][[1]]

  geneset.descriptions <- a[2][[1]]

  dd <- scan(filename, what = "", sep = "\t", quote = NULL)


  nn <- length(geneset.names)
  n <- length(dd)
  ox <- rep(NA, nn)

  ii <- 1
  for (i in 1:nn) {
    cat(i)
    while ((dd[ii] != geneset.names[i]) | (dd[ii + 1] != geneset.descriptions[i])) {
      ii <- ii + 1
    }
    ox[i] <- ii
    ii <- ii + 1
  }

  genesets <- vector("list", nn)

  for (i in 1:(nn - 1)) {
    cat(i, fill = T)
    i1 <- ox[i] + 2
    i2 <- ox[i + 1] - 1
    geneset.descriptions[i] <- dd[ox[i] + 1]
    genesets[[i]] <- dd[i1:i2]
  }

  geneset.descriptions[nn] <- dd[ox[nn] + 1]
  genesets[[nn]] <- dd[(ox[nn] + 2):n]
  out <- list(genesets = genesets, geneset.names = geneset.names, geneset.descriptions = geneset.descriptions)
  class(out) <- "GSA.genesets"
  return(out)
}

get_BC_eset <- function(pData) {
  # Format as EListRaw
  E <- raw_agilent[, pData$Sample_ID]
  E <- new("EListRaw", list(E = E, genes = data.frame(ProbeName = raw_agilent$ProbeID), targets = colnames(E), source = 'agilent.median'))
  
  # Quantile normalization
  eset.rma <- normalizeBetweenArrays(object = E, method = "quantile")
  
  # Average over quadruplicates
  eset.rma <- avereps(x = eset.rma, ID = E$genes$ProbeName) 
  
  # Batch correct
  # Export from EList to matrix
  eset.rma_m <- eset.rma$E
  
  # Subset for 38,387 known probes:
  eset.rma_m <- eset.rma_m[rownames(Annotation.known), ]
  
  batch <- pData$Batch
  modcombat <- model.matrix(~1, data = pData)
  eset.batch <- ComBat(dat = eset.rma_m, batch = batch, mod = modcombat, par.prior = TRUE, prior.plots = F)
  return(eset.batch)
}

sigboxplot <- function(data, include_celltypes = NULL, split.by = NA, plot_title = 'Estimated cell type composition differences', subtitle = '', color = c(), combs = NULL) {
  if (is.null(include_celltypes)) {
    include_celltypes <- colnames(data)[2:11]
  }
  plotdata <- melt(data = data, id = c(colnames(data)[!colnames(data) %in% include_celltypes]))
  ggplot(data = plotdata, aes(x = variable, y = value, fill = !! sym(split.by))) +
  geom_boxplot(aes(fill = !! sym(split.by)), position = position_dodge())  +
  scale_y_continuous(limits = c(0, 1)) + 
  stat_compare_means(comparisons = combs, mapping = aes(group = !! sym(split.by)), label = "p.signif", label.y = 0.9, hide.ns = T, method = 'wilcox.test') + 
  xlab(label = 'Cell type') + ylab(label = 'Proportion') + 
  ggtitle(label = plot_title, subtitle = subtitle) +
  theme_pubclean() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.key = element_rect(fill = 'NA')) + scale_fill_manual(values = color)
}

barplot.seurat <- function(obj, id1, id2){

    df.plot <- obj@meta.data %>% 
    group_by(.[[id1]]) %>%
    count(.[[id2]]) %>%
    mutate(Perc = n / sum(n) * 100) 

    colnames(df.plot)[1:2] <- c(id1, id2)

    ggplot(df.plot, aes_string(id1, "Perc", fill = id2)) + 
    geom_bar(stat = "Identity") + ylab("Percentage of cells") +
    guides(x = guide_axis(angle = 90)) + theme_minimal() 
}
```

### Read raw microarrays
```{r}
# Redo this, read in per Agilent-ID, select and order probes per agilent-ID. Only published data
agilent_platforms <- unique(pData_412$agilent_id)
path_published <- '//vf-DataSafe/DataSafe$/Div0/ict/optimized_SAGA_2457/GSE109391_RAW'
elists <- c()
for (platform in agilent_platforms) {
  pData <- pData_412[pData_412$agilent_id == platform & pData_412$Protocol == 'published', ]
  files <- pData$Filename
  raw_published <- read.maimages(path = path_published, files = files, source = "agilent.median", green.only = T, 
                                 columns = list(G = "gMedianSignal"), annotation = c("ProbeName"))
  
  colnames(raw_published) <- pData$Sample_ID
  # Subset for 39428 probes from Agilent ID026655
  index <- raw_published$genes$ProbeName %in% row.names(Annotation)
  raw_published <- raw_published[index, ]
  raw_published <- raw_published[order(raw_published$genes$ProbeName), ]
  elists <- append(elists, list(raw_published))
}

# New data
pData <- pData_412[!pData_412$Protocol == 'published', ]
files <- pData$Filename
path_new <- '//vf-DataSafe/DataSafe$/Div0/ict/optimized_SAGA_2457/SAGA_RAW/SAGA_RAW'
raw_published <- read.maimages(path = path_new, files = files, source = "agilent.median", green.only = T, 
                               columns = list(G = "gMedianSignal"), annotation = c("ProbeName"))

colnames(raw_published) <- pData$Sample_ID
# Subset for 39428 probes from Agilent ID026655
index <- raw_published$genes$ProbeName %in% row.names(Annotation)
raw_published <- raw_published[index, ]
raw_published <- raw_published[order(raw_published$genes$ProbeName), ]
elists <- append(elists, list(raw_published))

# Combine datasets
SAGA_RAW <- cbind.EList(elists[[1]], elists[[2]], elists[[3]], elists[[4]], elists[[5]], elists[[6]], elists[[7]])

# Export quadruplicates
SAGA_matrix <- data.frame(ProbeID = SAGA_RAW$genes$ProbeName, SAGA_RAW$E)
write.table(SAGA_matrix, file = "RAW_FullSagaSet412_quadruplicates_FINAL.txt", sep = "\t", row.names = TRUE, col.names = NA)
```


### Read quads and normalize and average 
```{r}
# Read quads
raw_agilent <- read.table('/exports/immu-staallab/Micah/RAW_FullSagaSet700_quadruplicates_FINAL.txt', sep = '\t', header = T)[, -1]

# Write data
write.table(eset.batch, file = "/exports/immu-staallab/Micah/ESET_RMA_COMBAT_FullSagaSet412_FINAL.txt", sep = "\t", row.names = TRUE, col.names = NA)

# How to read the data again
eset.batch <- read.table("/exports/immu-staallab/Micah/ESET_RMA_COMBAT_FullSagaSet412_FINAL.txt", sep = "\t", row.names = 1, header = T)
```


### Quantile normalization and averaging over quadruplicates
```{r}
# Quantile normalization
eset.rma <- normalizeBetweenArrays(object = raw_agilent, method = "quantile")

# Average over quadruplicates
eset.rma <- avereps(x = eset.rma, ID = eset.rma$genes$ProbeName)
```


### Batch correction using ComBat
```{r}
# Export from EList to matrix
eset.rma_m <- eset.rma$E

# Subset for 38,387 known probes:
eset.rma_m <- eset.rma_m[rownames(Annotation.known), ]

batch <- pData_final$Batch
modcombat <- model.matrix(~1, data = pData_final)
eset.batch <- ComBat(dat = eset.rma_m, batch = batch, mod = modcombat, par.prior = TRUE, prior.plots = F)
write.table(eset.batch, file = "/exports/immu-staallab/Micah/ESET_RMA_COMBAT_FullSagaSet412_FINAL.txt", sep = "\t", row.names = TRUE, col.names = NA)

# How to read data again
eset.batch <- read.table("/exports/immu-staallab/Micah/ESET_RMA_COMBAT_FullSagaSet700_FINAL.txt", sep = "\t", row.names = 1, header = T)
```

### Export final dataset
```{r}
# Export from EList to matrix
eset.rma_m <- eset.rma$E

# Subset for 38,387 known probes:
eset.rma_m <- eset.rma_m[rownames(Annotation.known), ]

# Export to csv file
write.csv(eset.rma_m, file = "/exports/immu-staallab/Micah/ESET_RMA_Final_258.txt")

# How to read data again
eset.rma <- read.csv(file = "/exports/immu-staallab/Micah/ESET_RMA_Final_258.txt", row.names = 1)
```


### Create mixture file for deconvolution
```{r}
# Export from EList to matrix
eset.rma_m <- eset.rma$E

# Subset for 38,387 known probes:
eset.rma_m <- eset.rma_m[rownames(Annotation.known), ]

# Select one probe for each gene (select probe with highest variation)
eset <- data.frame(eset.rma_m, GeneSymbol = Annotation.known$GeneSymbol_FINAL)
eset.sd <- apply(eset[1:length(eset) - 1], 1, sd)
eset.sd.order <- eset[order(eset.sd, decreasing = T), ]

# Add GeneSymbol column
eset.filter <- subset(eset.sd.order, !duplicated(eset.sd.order$GeneSymbol))
eset.filter <- relocate(eset.filter, "GeneSymbol")
eset.filter$GeneSymbol <- toupper(eset.filter$GeneSymbol)

# Export to file
write.table(eset.filter, "/exports/immu-staallab/Micah/mixture_batch_412_final.txt", quote = F, sep = "\t", row.names = F)
```




### t-SNE to show difference in transforming cultures
```{r}
library(Rtsne)

set.seed(37)
tsne_out <- Rtsne(t(eset.rma), dims = 2, perplexity = 16, theta = 0.5, pca = FALSE, max_iter = 1000, verbose = FALSE, is_distance = FALSE, check_duplicates = FALSE)
set.seed(37)
tsne_out_BC <- Rtsne(t(eset.batch), dims = 2, perplexity = 16, theta = 0.5, pca = FALSE, max_iter = 1000, verbose = FALSE, is_distance = FALSE, check_duplicates = FALSE)

set.seed(37)
tsne_out_BC <- Rtsne(t(predictors_batch), dims = 2, perplexity = 16, theta = 0.5, pca = FALSE, max_iter = 1000, verbose = FALSE, is_distance = FALSE, check_duplicates = FALSE)

set.seed(37)
tsne_out_BC <- Rtsne(matrix.unknown, dims = 2, perplexity = 1, theta = 0.5, pca = FALSE, max_iter = 1000, verbose = FALSE, is_distance = FALSE, check_duplicates = FALSE)
tsne_plot_BC <- data.frame(x = tsne_out_BC$Y[, 1], y = tsne_out_BC$Y[, 2])
tsne_plot_BC <- bind_cols(tsne_plot_BC, pData_test, predict = Prediction_SVM.Caret$Prediction.SVM.Caret)


tsne_plot <- data.frame(x = tsne_out$Y[, 1], y = tsne_out$Y[, 2])
tsne_plot <- bind_cols(tsne_plot, pData_final)

tsne_plot_BC <- data.frame(x = tsne_out_BC$Y[, 1], y = tsne_out_BC$Y[, 2])
tsne_plot_BC <- bind_cols(tsne_plot_BC, pData_final)

tsnes <- lapply(c("Batch", "clonal_evidence", "IVIM", "Class"), function(x) {
  p <- ggplot(tsne_plot) +
    geom_point(aes_string(x = "x", y = "y", color = x)) +
    ggpubr::theme_pubclean() +
    ggtitle(label = paste0(x)) +
    theme(legend.position = "right", legend.key = element_rect(fill = NA))
  if (x == 'clonal_evidence') {
    p <- p + scale_color_manual(values = clone_colors)
  }
  if (x == 'IVIM') {
    p <- p + scale_color_manual(values = ivim_colors)
  }
  if (x == 'Class') {
    p <- p + scale_color_manual(values = class_colors)
  }
  return(p)
})

wrap_plots(tsnes)

tsnes_bc <- lapply(c("Vector", "predict", "IVIM", "Design"), function(x) {
  p <- ggplot(tsne_plot_BC) +
    geom_point(aes_string(x = "x", y = "y", color = x)) +
    ggpubr::theme_pubclean() +
    ggtitle(label = paste0(x)) +
    theme(legend.position = "right", legend.key = element_rect(fill = NA))
  if (x == 'clonal_evidence') {
    p <- p + scale_color_manual(values = clone_colors)
  }
  if (x == 'IVIM') {
    p <- p + scale_color_manual(values = ivim_colors)
  }
  if (x == 'Class') {
    p <- p + scale_color_manual(values = class_colors)
  }
  if (x == 'Design') {
    p <- p + scale_color_manual(values = condition_colors)
  }
  return(p)
})

wrap_plots(tsnes_bc)
```

### PCA
```{r}
# Run PCA
pca <- prcomp(t(eset.rma$E))
pca <- prcomp(t(eset.batch))
pca <- prcomp(t(matrix.gsea))
# PCA on one batch
pca <- prcomp(t(eset.batch[, pData_model$Sample_ID[pData_model$Batch == 51],]))

# PCA testdata
pca <- prcomp(matrix.unknown)
pca_df <- data.frame(x = pca$x[, 1], y = pca$x[, 2]) %>% bind_cols(., pData_test)
pca_df$predict <- Prediction_SVM.Caret$Prediction.SVM.Caret

# PCA on predictors
predictors_batch <- eset.batch[GA.FINAL[['ga']][['final']]]

predictors_batch <- predictors_batch[!rownames(predictors_batch) == 'NA',]

pca <- prcomp(t(predictors_batch))

# Format results for plotting
pca_summary <- summary(pca)
pca_df <- data.frame(x = pca$x[, 1], y = pca$x[, 2]) %>% bind_cols(., pData_final)
pca_df <- data.frame(x = pca$x[, 1], y = pca$x[, 2]) %>% bind_cols(., pData_model)
pca_df <- data.frame(x = pca$x[, 1], y = pca$x[, 2]) %>% bind_cols(., pData_model[pData_model$Batch == 51, ])

# Single PC plot
ggplot(data = pca_df, aes_string("x", "y", col = "Design")) +
  geom_point() +
  ggpubr::theme_pubclean() +
  ggtitle(label = paste0("Design")) +
  theme(legend.position = "right", legend.key = element_rect(fill = NA)) + scale_color_manual(values = condition_colors) + xlab(label = paste0("Explained Variance (", round(pca_summary$importance[2, 1] * 100, 1), "%)")) + ylab(label = paste0("Explained Variance (", round(pca_summary$importance[2, 2] * 100, 1), "%)"))

# Create multiple PC plots colored by different properties
pcas <- lapply(c("Batch", "agilent_id", "IVIM", "Class"), function(x) {
  p <- ggplot(data = pca_df, aes_string("x", "y", col = x)) +
    geom_point() +
    ggpubr::theme_pubclean() +
    ggtitle(label = paste0(x)) +
    theme(legend.position = "right", legend.key = element_rect(fill = NA))
  if (x == 'clonal_evidence') {
    p <- p + scale_color_manual(values = clone_colors)
  }
  if (x == 'IVIM') {
    p <- p + scale_color_manual(values = ivim_colors)
  }
  if (x == 'Class') {
    p <- p + scale_color_manual(values = class_colors)
  }
  return(p)
})

# Wrap plots
wrap_plots(pcas) & xlab(label = paste0("Explained Variance (", round(pca_summary$importance[2, 1] * 100, 1), "%)")) & ylab(label = paste0("Explained Variance (", round(pca_summary$importance[2, 2] * 100, 1), "%)"))
```


### Find differentially expressed genes between samples - LIMMA
```{r}
setwd(dir = "~/SAGA_GEX_Analysis/saga258/")
# Create design matrix
targets_152 <- read.table('SAGA_GEX_Analysis/Published_data/SAGA_Targets_FINAL_152.txt', header = T, row.names = 1)

pData_412 <- read.csv('/home/mmallee/pData_train_412.csv')

pData <- pData_152
pData <- targets_152
pData <- pData_154
pData <- pData_258
pData <- pData_412
groups <- factor(pData$Design)
design <- model.matrix(~ 0 + groups)
rownames(design) <- pData$Sample
colnames(design) <- c("Mutagenic", "Mock", "Safe")

# Create contrast matrix
cont.matrix <- makeContrasts(
  c1 = Mutagenic - Mock,
  c2 = Safe - Mock,
  c3 = Mutagenic - (Mock + Safe) / 2,
  c4 = Mutagenic - Safe,
  levels = design
)

# Fit model and compute estimated coefficients and standard errors
fit <- lmFit(eset_258, design)
fit <- lmFit(eset_154, design)
fit <- lmFit(eset.batch, design)
fit <- lmFit(ok_152, design)
fit <- lmFit(eset_412, design)
fit <- contrasts.fit(fit, contrasts = cont.matrix)
fit2 <- eBayes(fit)


# Analyze results
## Output: Supplementary Data 2
library(writexl)
# Supplementary Data 2 tab 1
topMvMoS <- topTable(fit2, coef = 3, n = Inf, adjust.method = "BH", sort.by = "logFC")
topMvMoS <- cbind(topMvMoS, Annotation.known[row.names(topMvMoS), ])
write.table(topMvMoS, file = "/home/mmallee/SAGA_GEX_Analysis/saga412/Toplist_FullSagaSet412_Mutagenic_vs_Mock_and_Safe_FINAL_batch.txt", sep = "\t", col.names = NA)

# Supplementary Data 2 tab 3: transforming - mock
topMvMo <- topTable(fit2, coef = 1, n = Inf, adjust.method = "BH", sort.by = "logFC")
topMvMo <- cbind(topMvMo, Annotation.known[row.names(topMvMo), ])
write.table(topMvMo, file = "/home/mmallee/SAGA_GEX_Analysis/saga412/Toplist_FullSagaSet412_Mutagenic_vs_Mock_FINAL_batch.txt", sep = "\t", col.names = NA)

# Supplementary Data 2 tab 4: safe vs mock
topSvMo <- topTable(fit2, coef = 2, n = Inf, adjust.method = "BH", sort.by = "logFC")
topSvMo <- cbind(topSvMo, Annotation.known[row.names(topSvMo), ])
write.table(topSvMo, file = "/home/mmallee/SAGA_GEX_Analysis/saga412/Toplist_FullSagaSet412_Safe_vs_Mock_FINAL_batch.txt", sep = "\t", col.names = NA)

# Supplementary Data 2 tab 5
topMvS <- topTable(fit2, coef = 4, n = Inf, adjust.method = "BH", sort.by = "logFC")
topMvS <- cbind(topMvS, Annotation.known[row.names(topMvS), ])
write.table(topMvS, file = "/home/mmallee/SAGA_GEX_Analysis/saga412/Toplist_FullSagaSet412_Mutagenic_vs_Safe_FINAL_batch.txt", sep = "\t", col.names = NA)

sheets <- list("Mutagenic vs Mock and Safe" = topMvMoS, "Mutagenic vs Mock" = topMvMo, 
               'Mutagenic vs Safe' = topMvS, 'Safe vs Mock' = topSvMo)
# write_xlsx(sheets, "/home/mmallee/SAGA_GEX_Analysis/saga412/Supplemental_table_S2.xlsx")
write_xlsx(sheets, "/home/mmallee/SAGA_GEX_Analysis/saga258/Supplemental_table_S2b.xlsx")



# Read again
topMvMo <- read.table('~/SAGA_GEX_Analysis/saga258/Toplist_FullSagaSet258_Mutagenic_vs_Mock_FINAL_batch.txt', header = T, sep = '\t')
top.1_orig_MvMo <- readxl::read_excel(path = "~/Hannover_data/1-s2.0-S1525001621003233-mmc3.xlsx", sheet = 3, col_names = T, skip = 1)

# Original vs new volcano plots
EnhancedVolcano(toptable = topMvS, x = "logFC", y = "P.Value", lab = topMvS$GeneSymbol_FINAL, title = "Mutagenic vs Safe", subtitle = "Unpublished", xlim = c(-2, 4), ylim = c(-1, 40), pCutoff = 0.05)
EnhancedVolcano(toptable = topMvMo, x = "logFC", y = "P.Value", lab = topMvMo$GeneSymbol_FINAL, title = "Mutagenic vs Mock", subtitle = "Unpublished", xlim = c(-2, 4), ylim = c(-1, 40), pCutoff = 0.05)
EnhancedVolcano(toptable = topMvMoS, x = "logFC", y = "P.Value", lab = topMvMoS$GeneSymbol_FINAL, title = "Mutagenic vs Safe+Mock", subtitle = "Unpublished", xlim = c(-2, 4), ylim = c(-1, 40), pCutoff = 0.05)
EnhancedVolcano(toptable = topSvMo, x = "logFC", y = "P.Value", lab = topSvMo$GeneSymbol_FINAL, title = "Safe vs Mock", subtitle = "Unpublished", xlim = c(-2, 4), ylim = c(-1, 40), pCutoff = 0.05)

EnhancedVolcano(toptable = top.2_orig_SvMo, x = "logFC", y = "P.Value", lab = top.2_orig_SvMo$GeneSymbol_FINAL, title = "Safe vs Mock", subtitle = "Published", xlim = c(-2, 4), ylim = c(-1, 40), pCutoff = 0.05)

#look at overlapping genes

overlapping_deg_MvS_up <- intersect(topMvS$GeneSymbol_FINAL[topMvS$logFC > 1], top.4_origMvS$GeneSymbol_FINAL[top.4_origMvS$logFC > 1])
overlapping_deg_MvS_down <- intersect(topMvS$GeneSymbol_FINAL[topMvS$logFC < -1], top.4_origMvS$GeneSymbol_FINAL[top.4_origMvS$logFC < -1])

overlapping_deg_MvMoS_up <- intersect(topMvMoS$GeneSymbol_FINAL[topMvMoS$logFC > 1], top.3_orig_MvMoS$GeneSymbol_FINAL[top.3_orig_MvMoS$logFC > 1])
overlapping_deg_MvMoS_down <- intersect(topMvMoS$GeneSymbol_FINAL[topMvMoS$logFC < -1], top.3_orig_MvMoS$GeneSymbol_FINAL[top.3_orig_MvMoS$logFC < -1])

overlapping_deg_MvMo_up <- intersect(topMvMo$GeneSymbol_FINAL[topMvMo$logFC > 1], top.1_orig_MvMo$GeneSymbol_FINAL[top.1_orig_MvMo$logFC > 1])
overlapping_deg_MvMo_down <- intersect(topMvMo$GeneSymbol_FINAL[topMvMo$logFC < -1], top.1_orig_MvMo$GeneSymbol_FINAL[top.1_orig_MvMo$logFC < -1])

overlapping_deg_SvMo_up <- intersect(topSvMo$GeneSymbol_FINAL[topSvMo$logFC > 1], top.2_orig_SvMo$GeneSymbol_FINAL[top.2_orig_SvMo$logFC > 1])
overlapping_deg_SvMo_down <- intersect(topSvMo$GeneSymbol_FINAL[topSvMo$logFC < -1], top.2_orig_SvMo$GeneSymbol_FINAL[top.2_orig_SvMo$logFC < -1])
```

### Compare original DEGs vs new
```{r}
# Select matching probes
matching <- intersect(top.1_orig_MvMo$PROBEID, rownames(topMvMo))

degdf <- data.frame(probe = top.1_orig_MvMo$PROBEID, Original_logFC = top.1_orig_MvMo$logFC)
degdf <- data.frame(degdf, New_logFC = topMvMo[match(degdf$probe, rownames(topMvMo)), 1])
ggscatter(data = degdf, x = 'Original_logFC', y = 'New_logFC', add = 'reg.line', add.params = list(color = 'red', fill = 'lightgray'), conf.int = T) + stat_cor(method = 'pearson') + ggtitle(label = 'logFC correlation for matching probes between published data \nDEGs and unpublished data DEGs', subtitle = 'Mutagenic vs. Mock')
```



### GSEA with fgsea
```{r}
# BiocManager::install("fgsea")
library(fgsea)

topMvMo <- read.table('~/SAGA_GEX_Analysis/saga258/Toplist_FullSagaSet258_Mutagenic_vs_Mock_FINAL_batch.txt', header = T, sep = '\t')
topMvMoS_412 <- read.table('~/SAGA_GEX_Analysis/saga412/Toplist_FullSagaSet412_Mutagenic_vs_Mock_and_Safe_FINAL_batch.txt', header = T, sep = '\t')
# Use handselected first
sets <- readxl::read_excel(path = "~/Hannover_data/1-s2.0-S1525001621003233-mmc4.xlsx", sheet = 1, skip = 1, col_names = T)[, -2] # 106 GeneSets
# sets_MvsMo_orig <- readxl::read_excel(path = "~/Hannover_data/1-s2.0-S1525001621003233-mmc4.xlsx", sheet = 2, skip = 1, col_names = T)[, -2] # 106 GeneSets
# colnames(sets_MvsMo_orig)[c(1, 4, 5)] <- c('pathway', 'pval', 'padj')

sets <- lapply(split(sets[, -1], sets[, 1]), unlist)
sets <- lapply(sets, function(x) {
  x[!is.na(x)]
})

mast_sets <- GSA.read.gmt(filename = '~/genesets.v2023.2.Hs_mast.gmt')
names(mast_sets$genesets) <- mast_sets$geneset.names
sets <- append(sets, mast_sets$genesets)



genes <- unique(unlist(sets))


## Add signature matrix gene sets
# sig_mat <- read.table('/exports/immu-staallab/Micah/signature_matrix_DMC_C.txt', header = T, sep = '\t', row.names = 1)
# 
# sig_mat$highest <- colnames(sig_mat)[max.col(sig_mat, ties.method = "first")]
# table(sig_mat$highest)
# 
# sig_mat_sets <- lapply(unique(sig_mat$highest), function(ct) {
#   tmp <- sig_mat[sig_mat$highest == ct, ]
#   return(rownames(tmp))
# })
# names(sig_mat_sets) <- paste0(colnames(sig_mat[1:10]), '_SIG_MAT')
# 
# sets <- append(sets, sig_mat_sets)

eset.batch <- eset_412

# Make DataSet with one probe for each gene (select probe with highest variation)
eset <- data.frame(eset.batch, GeneSymbol = Annotation.known$GeneSymbol_FINAL)
eset.gsea <- subset(eset, toupper(eset$GeneSymbol) %in% genes) # take all probes with entries in GSEA-Sets
eset.sd <- apply(eset.gsea[1:length(eset.gsea) - 1], 1, sd)
eset.sd.order <- eset.gsea[order(eset.sd, decreasing = T), ]
eset.gsea <- subset(eset.sd.order, !duplicated(eset.sd.order$GeneSymbol))
matrix.gsea <- as.matrix(eset.gsea[, c(1:length(eset.gsea) - 1)])
Annotation.GSEA <- Annotation.known[row.names(eset.gsea), ]

# Filter for unique genes
genes <- genes[genes %in% toupper(Annotation.GSEA$GeneSymbol_FINAL)]

topMvMoS_412 <- topMvMoS_412[topMvMoS_412$X %in% rownames(Annotation.GSEA), ]



# Creating ranks
ranks <- topMvMoS_412$logFC[toupper(topMvMoS_412$GeneSymbol_FINAL) %in% genes]
names(ranks) <- toupper(topMvMoS_412$GeneSymbol_FINAL[toupper(topMvMoS_412$GeneSymbol_FINAL) %in% genes])

# top.1_orig <- readxl::read_excel(path = "~/Hannover_data/1-s2.0-S1525001621003233-mmc3.xlsx", sheet = 3, col_names = T, skip = 1)
# top.1_orig <- top.1_orig[top.1_orig$PROBEID %in% rownames(Annotation.GSEA), ]
# ranks <- top.1_orig$logFC[toupper(top.1_orig$GeneSymbol_FINAL) %in% genes]
# names(ranks) <- toupper(top.1_orig$GeneSymbol_FINAL[toupper(top.1_orig$GeneSymbol_FINAL) %in% genes])

# Plot ranks
barplot(sort(ranks, decreasing = T))


# Load pathways (GeneSets)
pathways <- gmtPathways('/home/mmallee/SAGA_GEX_Analysis/Full_dataset/c2.c3.c5.c6.h.6.2.symbols.gmt')

fgseaRes <- fgsea(pathways = sets, ranks, minSize = 15, maxSize = 500)
fgseaRes <- fgsea::fgseaMultilevel(pathways = sets, ranks, minSize = 15, maxSize = 500, nPermSimple = 10000)

fgseaRes_full <- fgsea::fgseaMultilevel(pathways = pathways, ranks, minSize = 15, maxSize = 500, nPermSimple = 10000)
# Plot NES
plotEnrichment(sets[["WONG_EMBRYONIC_STEM_CELL_CORE"]], ranks)

fgseaRes$color <- sets_MvsMo_orig$`Color (Hex)`[match(toupper(fgseaRes$pathway), sets_MvsMo_orig$`GENE SET`)]

plot(log10(fgseaRes$padj), fgseaRes$NES, ylab = "Normalized enrichment score (NES)", xlab = "FDR (log10)", pch = 16, cex = 3)
segments(log10(0.10),-5,log10(0.10),5, lty = 2)

plot(log10(sets_MvsMo_orig$padj), y = sets_MvsMo_orig$NES, ylab = "Normalized enrichment score (NES)", xlab = "FDR (log10)", pch = 16, cex = 3, col = sets_MvsMo_orig$`Color (Hex)`, ylim = c(-2, 5))
segments(log10(0.10),-5,log10(0.10),5, lty = 2)

topPathwaysUp <- fgseaRes[NES > 0][head(order(pval), n=10), pathway]
topPathwaysDown <- fgseaRes[NES < 0][head(order(pval), n=10), pathway]
topPathways <- c(topPathwaysUp, rev(topPathwaysDown))

sign_genes <- fgseaRes[grep("SIG_MAT", fgseaRes$pathway), ][order(NES, decreasing = T)]

plotGseaTable(sets[topPathways], ranks, fgseaRes, gseaParam=0.5)
plotGseaTable(sets[sign_genes$pathway], ranks, fgseaRes, gseaParam=0.5)



# Export gene sets
setsdf <- as.data.frame(cbind(sets))
mx <- max(lengths(sets))
d1 <- as.data.frame(t(data.frame(lapply(sets, `length<-`, mx))))
d1 <- cbind(GeneSet = rownames(d1), d1)

writexl::write_xlsx(d1, 'Supplemental_Table_S3_GSEA_Sets.xlsx')
```

### Remove mast specific genes and export expression matrix
```{r}
# Load prefiltered probes
iqr_filt <- read.table('~/SAGA_Model_3.0/Annotation.matrix.train.FINAL_batch.txt', header = T)
GA.FINAL <- readRDS("~/SAGA_Model_3.0/gafinal.rds")
predictors <- Annotation.known[GA.FINAL$optVariables, ]


# Look at signature matrix genes
sig_mat <- read.table('/exports/immu-staallab/Micah/signature_matrix_DMC_C.txt', header = T, row.names = 1)
sig_mat$highest <- colnames(sig_mat)[max.col(sig_mat, ties.method = "first")]
table(sig_mat$highest)
mast_genes <- rownames(sig_mat[sig_mat$highest == 'Mast_cell', ])

# Mast genes in prefilt and predictors
mast_filt <- mast_genes[mast_genes %in% toupper(iqr_filt$GeneSymbol_FINAL)]
mast_pred <- mast_genes[mast_genes %in% toupper(predictors$GeneSymbol_FINAL)]
FindVariableFeatures()


# Signature genes in prefilt and predictors
sig_filt <- rownames(sig_mat)[rownames(sig_mat) %in% toupper(iqr_filt$GeneSymbol_FINAL)] #%>% length()
sig_pred <- rownames(sig_mat)[rownames(sig_mat) %in% toupper(predictors$GeneSymbol_FINAL)] #%>% length()

ct_pseudo <- AggregateExpression(real_seurat, group.by = c('cell_type', 'Condition'), assays = 'RNA', return.seurat = T)

mock_cells <- subset(real_seurat, subset = Condition == 'mock')
mutagenic_cells <- subset(real_seurat, subset = Condition == 'RSF91')

mock_hvf <- HVFInfo(object = mock_cells)
mutagenic_hvf <- HVFInfo(object = mutagenic_cells)
pseudo_hvf <- HVFInfo(object = ct_pseudo, assay = 'RNA', method = 'vst')


total_genes <- unique(c(rownames(sig_mat), toupper(names(iqr_scaled_filt))))


# Look at variable scRNAseq genes
real_seurat <- readRDS('/exports/immu-staallab/Micah/pData.final.RDS')
Idents(real_seurat) <- 'cell.types'
new.cluster.ids <- c("Mast_cell", "Basophil", "Mast_cell", "Mast_cell", "Mast_cell", "Mast_cell/Basophil", 
    "Macrophage", "Basophil", "Macrophage", 'Macrophage', 'Mast_cell', 'Macrophage', 'DC/Macrophage', 'HSPC', 'Mast_cell', 'Unknown', 'CMP', 'Neutrophil', 'MEP', 'Unknown')
names(new.cluster.ids) <- levels(real_seurat)
real_seurat <- RenameIdents(real_seurat, new.cluster.ids)
real_seurat$cell_type <- real_seurat@active.ident


iqr_scaled <- apply(real_seurat@assays[["RNA"]]@counts, 1, IQR)
boxplot(iqr_scaled)
iqr_scaled_filt <- iqr_scaled[iqr_scaled > 0.2]

# scRNA filt genes in prefilt and predictors
scRNA_genes_iqr0.2 <- toupper(names(iqr_scaled_filt))[toupper(names(iqr_scaled_filt)) %in% toupper(iqr_filt$GeneSymbol_FINAL)]

scRNA_genes_iqr0.2 %in% toupper(predictors$GeneSymbol_FINAL)  %>% sum()

seq_filt <- scRNA_genes_iqr0.2
seq_pred <- scRNA_genes_iqr0.2[scRNA_genes_iqr0.2 %in% toupper(predictors$GeneSymbol_FINAL)]

# scRNA filt genes in signature matrix
scRNA_genes_iqr0.2 %in% rownames(sig_mat) %>% sum()



# Look at variable features
sum(toupper(real_seurat@assays[["RNA"]]@var.features) %in% rownames(sig_mat))
```


### Remove cell type specific genes and export expression matrix to build model on
```{r}
# Use a coefficient of variance cut-off = sd/mean
# Get pseudobulks per cell type
ct_pseudo <- AggregateExpression(object = real_seurat, assays = 'RNA', return.seurat = T, group.by = 'cell_type')


df <- data.frame(sd = apply(ct_pseudo@assays$RNA$counts, 1, sd), mean = apply(ct_pseudo@assays$RNA$counts, 1, mean))
df$CV <- df$sd/df$mean
df$CV[df$mean == 0] <- 0

# First select genes based on total counts, to not have only genes with < 50 counts
counts <- ct_pseudo@assays$RNA$counts

df$counts <- apply(counts, 1, sum)

ggplot(df, aes(x=log2(counts))) + geom_density() + geom_vline(xintercept = log2(50), color = '#CB181D') + ggtitle('Counts density \n(log2 transformed)') + ggplot(df, aes(x=CV)) + geom_density() + geom_vline(xintercept = 2, color = '#CB181D') + ggtitle('Coefficient of variance')# + geom_vline(xintercept = 3, linetype = 'dashed')

filt_genes <- rownames(df[df$counts >= 50 & df$CV > 2,])

sum(filt_genes %in% iqr_filt$GeneSymbol_FINAL) # 451 genes are present in predictor set
sum(filt_genes %in% predictors$GeneSymbol_FINAL) # 40 genes are present in predictor set

filt_probes <- rownames(Annotation.known[Annotation.known$GeneSymbol_FINAL %in% filt_genes, ])

sum(filt_probes %in% iqr_filt$X) # 451 probes are present in predictor set
sum(filt_probes %in% rownames(predictors)) # 40 genes are present in predictor set

write.table(filt_probes, '/exports/immu-staallab/Micah/CV_filtered_probes.txt')

ok <- read.table('/exports/immu-staallab/Micah/CV_filtered_probes.txt')
```


### Test simulated bulks for prediction
```{r}
library(SimBu)
library(Seurat)

dataset <- dataset(annotation = real_seurat@meta.data, count_matrix = real_seurat@assays$RNA$counts, name = "test_dataset")
simulation_random <- simulate_bulk(data = dataset, scenario = "random", scaling_factor = "NONE")

real_seurat$ID <- colnames(real_seurat)
mock_cells <- subset(real_seurat, subset = Condition == 'mock')
mutagenic_cells <- subset(real_seurat, subset = Condition == 'RSF91')

dataset_mock <- SimBu::dataset(annotation = mock_cells@meta.data, count_matrix = mock_cells@assays$RNA$counts, name = 'Mock', filter_genes = F)
dataset_mutagenic <- SimBu::dataset(annotation = mutagenic_cells@meta.data, count_matrix = mutagenic_cells@assays$RNA$counts, name = 'Mutagenic', filter_genes = F)

make_pseudobulks <- function(seurat_obj) {
  weighted_sims_mast <- lapply(c(0, 0.2, 0.4, 0.6, 0.8, 0.99), function(x){
    simulate_bulk(data = seurat_obj, scenario = "weighted", scaling_factor = "NONE", weighted_cell_type = 'Mast_cell', weighted_amount = x, nsamples = 20)})
  weighted_sims_mast_bas <- lapply(c(0, 0.2, 0.4, 0.6, 0.8, 0.99), function(x){
    simulate_bulk(data = seurat_obj, scenario = "weighted", scaling_factor = "NONE", weighted_cell_type = 'Mast_cell/Basophil', weighted_amount = x, nsamples = 20)})
    weighted_sims_bas <- lapply(c(0, 0.2, 0.4, 0.6, 0.8, 0.99), function(x){
    simulate_bulk(data = seurat_obj, scenario = "weighted", scaling_factor = "NONE", weighted_cell_type = 'Basophil', weighted_amount = x, nsamples = 20)})
  merged <- merge_simulations(c(weighted_sims_mast, weighted_sims_mast_bas, weighted_sims_bas))
  return(merged)
}

simulated_bulks <- lapply(list(dataset_mock, dataset_mutagenic), make_pseudobulks)

simulated_bulks_merged <- merge_simulations(simulated_bulks)
plot_simulation(simulated_bulks_merged)

bulk_counts <- as.data.frame(simulated_bulks_merged[["bulk"]]@assays@data@listData[["bulk_counts"]])



####### PURE PSEUDOBULKS
ct_pseudo <- AggregateExpression(real_seurat, group.by = 'Sample', assays = 'RNA')
bulk_counts <- as.data.frame(ct_pseudo$RNA)



###### end pure pseudobulks

# First load eset.batch and create matrix with only one probe per gene above
# Make DataSet with one probe for each gene (select probe with highest variation)
eset <- data.frame(eset.batch, GeneSymbol = Annotation.known$GeneSymbol_FINAL)
eset.sd <- apply(eset[1:length(eset) - 1], 1, sd)
eset.sd.order <- eset[order(eset.sd, decreasing = T), ]
eset <- subset(eset.sd.order, !duplicated(eset.sd.order$GeneSymbol))
matrix.gsea <- as.matrix(eset[, c(1:length(eset) - 1)])
Annotation.GSEA <- Annotation.known[row.names(eset), ]

# Give simulated bulks correct probes as rownames
bulk_counts_matching_probes <- bulk_counts[rownames(bulk_counts) %in% Annotation.GSEA$GeneSymbol_FINAL, ]
rownames(bulk_counts_matching_probes) <- rownames(bulk_counts_matching_probes) <- rownames(Annotation.GSEA)[match(rownames(bulk_counts_matching_probes), Annotation.GSEA$GeneSymbol_FINAL)] %>% .[!is.na(.)]

# Predict
pData_simbu <- data_frame(Sample = ct_pseudo[["RNA"]]@Dimnames[[2]], Batch = rep(200, ncol(bulk_counts_matching_probes)), Design = c(rep(times =3, x = 'A2_Mock'), rep(times = 3, x = 'A1_Mutagenic')))

pData_simbu <- data.frame(Sample = simulated_bulks_merged[["bulk"]]@colData@rownames, Batch = rep(200, ncol(bulk_counts_matching_probes)), Design = c(rep('A2_Mock', 360), rep('A1_Mutagenic', 360)))

Prediction_SVM.Caret <- cbind(simulated_bulks_merged[["cell_fractions"]], Prediction_SVM.Caret, design = pData_simbu$Design)

ggscatter(Prediction_SVM.Caret, x= 'Mast_cell', y = 'transforming', add = 'reg.line', add.params = list(color = 'blue', fill = 'lightgray'), conf.int = T) + stat_cor(method = 'pearson', label.x = 0.5, color = 'red', label.y = 0.85) + facet_wrap(~design)

ggscatter(Prediction_SVM.Caret, x= 'Basophil', y = 'transforming', add = 'reg.line', add.params = list(color = 'blue', fill = 'lightgray'), conf.int = T) + stat_cor(method = 'pearson', label.x = 0.5, color = 'red', label.y = 0.85) + facet_wrap(~design)

colnames(Prediction_SVM.Caret)[5] <- 'mast_basophil'

ggscatter(Prediction_SVM.Caret, x= 'mast_basophil', y = 'transforming', add = 'reg.line', add.params = list(color = 'blue', fill = 'lightgray'), conf.int = T) + stat_cor(method = 'pearson', label.x = 0.5, color = 'red', label.y = 0.85) + facet_wrap(~design)
```


### SAGA-GSEA
```{r}
# Load packages
library(phenoTest)
library(limma)
library(gridExtra)
library(pROC)
library(caret)
library(purrr)

# Read data
## load the Top 11 SAGA features from SVM-GA for FullSet230: 
genefilt_set <- read.delim("~/SAGA_Model_4.0_nomastgenes/FinalSet230_optVars_GeneticAlgorithm.txt", header = T, sep = "\t", row.names = 1)
genefilt_set <- rownames(genefilt_set)
gafinal <- readRDS("~/SAGA_Model_5.0_Full_dataset/Final_412/gafinal.rds")
new_set <- gafinal$optVariables
sets_11 <- rownames(saga::Top11)
topMvMo <- read.table('~/SAGA_GEX_Analysis/saga258/Toplist_FullSagaSet258_Mutagenic_vs_Mock_FINAL_batch.txt', header = T, sep = '\t')
topMvMo_412 <- read.table('~/SAGA_GEX_Analysis/saga412/Toplist_FullSagaSet412_Mutagenic_vs_Mock_FINAL_batch.txt', header = T, sep = '\t')
MvMo_114 <- topMvMo[topMvMo$X %in% genefilt_set,]
upreg_99 <- MvMo_114[MvMo_114$logFC > 0, "X"]
fc_geneset <- read.table('fc_geneset104.txt')[,1]



MvMo_74 <- topMvMo_412[topMvMo_412$X %in% new_set,]
upreg_67 <- MvMo_74[MvMo_74$logFC > 0, "X"]

SAGA.CORE <- list(top114 = genefilt_set, top11 = sets_11, top99 = upreg_99)
SAGA.CORE <- list(top67 = upreg_67, top11 = sets_11)

# Read sample information and loop over batches seperately (each batch has mock samples available)
SIF <- read.delim("~/pData_train_258.csv", row.names = 1, header = TRUE, sep = "\t", stringsAsFactors = F)
SIF <- read.delim('~/Full_Sample_information_merged.csv', row.names = 1, header = T, sep = '\t')
SIF <- SIF[SIF$Day == 15,]

SIF <- read.csv('/home/mmallee/pData_train_412.csv', header = T)
batches <- levels(as.factor(SIF$Batch)) # how many assays / batches 
result.all <- NULL # dummy data.frame
SIF$Group[SIF$Design == 'A2_Mock'] <- 1

# Which batches have no mock samples?
table(SIF$Batch, SIF$Design) #  batch 11 has no mock, remove this
batches <- batches[-which(batches == 11)]

# Give each sample group nr 2:nsamples per batch
for (batch in batches) {
  nonmock <- SIF[SIF$Batch == batch & !SIF$Design == 'A2_Mock',]
  SIF[rownames(nonmock), "Group"] <- 2:(nrow(nonmock)+1)
}


# read quadruplicates for all day 15 samples
quad700 <- read.table('/exports/immu-staallab/Micah/RAW_FullSagaSet700_quadruplicates_FINAL.txt', sep = '\t', header = T)[, -1]
FC_100_all <- list()
FC_full <- list()
# SAGA-GSEA Loop over all batches 
# norm, average, make expressionset, make epheno object, run gsea, export results
for(batch in batches) {   
  # Subset quadruplicates for batch
  SIF.i <- SIF[SIF$Batch==batch, ]
  
  # Format as EListRaw so Limma functions perform as expected (log transform before normalization and averaging)
  E.batch <- quad700[, SIF.i$Sample_ID]
  E <- new("EListRaw", list(E = E.batch, genes = data.frame(ProbeName = quad700$ProbeID), targets = colnames(E.batch), source = 'agilent.median'))
  
  # Quantile normalization and averaging
  eset.rma <- normalizeBetweenArrays(object = E, method = "quantile")
  eset.rma <- avereps(eset.rma, ID = eset.rma$genes$ProbeName)
  matrix.gsea <- eset.rma$E
  
  # Create ExpressionSet
  metadata  <- data.frame(labelDescription = rep(NA, dim(SIF.i)[2]), row.names = colnames(SIF.i))
  phenoData <- new("AnnotatedDataFrame", data = SIF.i, varMetadata = metadata)
  rownames(phenoData) <- colnames(matrix.gsea)
  eset.gsea <- ExpressionSet(assayData = matrix.gsea, phenoData = phenoData)
  
  # Create ePheno object
  vars2test   <- list(ordinal = "Group") # Variables (here: Vectors) to test against MOCK, which is always = 1 in the SIF 
  epheno.gsea <- ExpressionPhenoTest(eset.gsea, vars2test, p.adjust.method = 'BH')
  
  # GSEA
  SAGA.GSEA <- gsea(x = epheno.gsea, gsets = SAGA.CORE , B = 6000, center = TRUE, test = "perm", p.adjust.method = 'BH')
  result <- summary.gseaData(SAGA.GSEA)[, c(1, 2, 3, 5, 8)] # extract results (only NES-normalized enrichment scores)
  
  
  # Export high mutagenic FC genes
  mutagenic_groups <- SIF.i$Group[SIF.i$Design == 'A1_Mutagenic']
  # Use lapply to iterate over mutagenic_groups
  FC_100 <- unlist(lapply(mutagenic_groups, function(group) {
    FC_100_mutagenic <- sort(SAGA.GSEA[["simulations"]][[paste0("Group.", group, '.fc')]][["fc.hr"]], decreasing = TRUE)[1:100]
  }))
  
  nrgroups <- unique(SIF.i$Group[!SIF.i$Group == 1])
  FC_per_set <- lapply(nrgroups, function(group) {
    fc_114 <- SAGA.GSEA[["simulations"]][[paste0("Group.", group, '.fc')]][["fc.hr"]][SAGA.CORE$top114]
    fc_11 <- SAGA.GSEA[["simulations"]][[paste0("Group.", group, '.fc')]][["fc.hr"]][SAGA.CORE$top11]
    fc_99 <- SAGA.GSEA[["simulations"]][[paste0("Group.", group, '.fc')]][["fc.hr"]][SAGA.CORE$top99]
    return(list(fc_114, fc_11, fc_99))
  })
  
  names(FC_per_set) <- SIF.i$Sample_ID[!SIF.i$Group == 1]
  
  FC_full <- append(FC_full, FC_per_set)
  FC_100_all <- append(FC_100_all, FC_100)
  
  # Export
  Group <- NULL    ### pull out the Group index number from the result table                  
  for (a in 1:nrow(result)) {Group[a] <- unlist(strsplit(as.character(result$variable[a]), '.', fixed = TRUE))[2] }
  result$Group     <- Group
  
  SIF.sub           <- SIF.i[SIF.i$Group != 1, c('Group','Design')] # pull out info of tested vectors                    
  SIF.sub$SampleID  <- row.names(SIF.sub)
  result.m          <- merge(SIF.sub,result, by.x="Group", by.y = "Group") # merge result with SIF for SampleIDs and FileNames
  result.all        <- rbind(result.all, result.m)                           # merge all results to one file
  write.table(result.m, file = paste("~/SAGA_GSEA_results/Results_SAGA.GSEA_Batch_", batch,".txt", sep = ""), sep = "\t", row.names = FALSE)
  # make pdf
  pdf(file=paste("~/SAGA_GSEA_results/SAGA.GSEA_Batch_", batch, ".pdf", sep = ""), useDingbats = F, width = 10, height = 10)
  grid.table(result.m, rows = NULL)
  plot.gseaData(SAGA.GSEA, es.nes ='nes', selGsets='top11')
  plot.gseaData(SAGA.GSEA, es.nes ='nes', selGsets='top67', selVars = 'Group.2.fc')
  # plot.gseaData(SAGA.GSEA, es.nes ='nes', selGsets='top114')
  # plot.gseaData(SAGA.GSEA, es.nes ='nes', selGsets='top99')
  # plot.gseaData(SAGA.GSEA, es.nes ='nes', selGsets='fc_104')
  dev.off()
}

saveRDS(FC_100_all, 'FC_100_all.rds')
saveRDS(FC_full, 'FC_full.rds')

# Read results again

results_true <- lapply(batches, function(batch){
  read.table(file = paste("~/SAGA_GSEA_results/Results_SAGA.GSEA_Batch_", batch,".txt", sep = ""), sep = "\t", header = T)
})
results_all <- data.table::rbindlist(results_true)

results_top11 <- results_all[results_all$geneSet == 'top11',]
results_top67 <- results_all[results_all$geneSet == 'top67',]
results_top114 <- results_all[results_all$geneSet == 'top114',]
# results_top159 <- results_all[results_all$geneSet == 'top159',]
results_top99 <- results_all[results_all$geneSet == 'top99',]

Results_11 <- merge(SIF, results_top11, by.x = 0, by.y = "SampleID", all.x = F)
Results_67 <- merge(SIF, results_top67, by.x = 0, by.y = "SampleID", all.x = F)
Results_114 <- merge(SIF, results_top114, by.x = 0, by.y = "SampleID", all.x = F)
# Results_159 <- merge(SIF, results_top159, by.x = 0, by.y = "SampleID", all.x = F)
Results_99 <- merge(SIF, results_top99, by.x = 0, by.y = "SampleID", all.x = F)
rownames(Results_11) <- Results_11$Row.names
rownames(Results_67) <- Results_67$Row.names
rownames(Results_114) <- Results_114$Row.names
# rownames(Results_159) <- Results_159$Row.names
rownames(Results_99) <- Results_99$Row.names


rm(results_top11, results_top114, results_top159, results_true, results_all)

roc.11 <- roc(Results_11$Class,                    
                  Results_11$nes,             
                  percent = TRUE, levels=c("untransforming","transforming"),
                  plot = T, auc.polygon = F, max.auc.polygon = F, col = "#686868", lwd = 4, grid = F, add = F,
                  print.auc = T, print.thres = "best")

roc.67 <- roc(Results_67$Class,                    
                  Results_67$nes,             
                  percent = TRUE, levels=c("untransforming","transforming"),
                  plot = T, auc.polygon = F, max.auc.polygon = F, col = "#686868", lwd = 4, grid = F, add = F,
                  print.auc = T, print.thres = "best")


# roc.114 <- roc(Results_114$Class,                    
#                   Results_114$nes,             
#                   percent = TRUE, levels=c("untransforming","transforming"),
#                   plot = T, auc.polygon = F, max.auc.polygon = F, col = "#686868", lwd = 4, grid = F, add = F,
#                   print.auc = T, print.thres = "best")   # best threshold = 1.0  ==> 88% sensitivity, 74% specificity

# roc.159 <- roc(Results_159$Class,
#                   Results_159$nes,             
#                   percent = TRUE, levels=c("untransforming","transforming"),
#                   plot = T, auc.polygon = F, max.auc.polygon = F, col = "#686868", lwd = 4, grid = F, add = F,
#                   print.auc = T, print.thres = "best")   # best threshold = 1.0  ==> 88% sensitivity, 74% specificity

# roc.99 <- roc(Results_99$Class,                    
#                   Results_99$nes,             
#                   percent = TRUE, levels=c("untransforming","transforming"),
#                   plot = T, auc.polygon = F, max.auc.polygon = F, col = "#686868", lwd = 4, grid = F, add = F,
#                   print.auc = T, print.thres = "best")   # best threshold = 1.0  ==> 88% sensitivity, 74% specificity


# roc.list <- list(roc.11, roc.114, roc.159)
roc.list <- list(roc.11, roc.114, roc.99)
roc.list <- list(roc.11, roc.67)

# extract auc
roc.list %>% 
  map(~tibble(AUC = .x$auc)) %>% data.table::rbindlist() -> data.auc
# data.auc$name <- c('Original predictors (11)', 'Cell type filtered predictors (114)', 'No gene filtering (159)')
# data.auc$name <- c('Original predictors (11)', 'Cell type filtered predictors (114)', 'Only upregulated (99)')
data.auc$name <- c('Original predictors (11)', 'Upregulated new predictors (67)')

# generate labels labels
data.auc %>% 
  mutate(label_long=paste0(name," , AUC = ",paste(round(AUC,2))),
         label_AUC=paste0("AUC = ",paste(round(AUC,2)))) -> data.labels


data.labels <- cbind(data.labels, lapply(roc.list, coords, 'best', input = 'threshold', ret = c("threshold", 'specificity', 'sensitivity')) %>% data.table::rbindlist())

# names(roc.list) <- c('Original predictors (11)', 'Cell type filtered predictors (114)', 'No gene filtering (159)')
# names(roc.list) <- c('Original predictors (11)', 'Cell type filtered predictors (114)', 'Only upregulated (99)')
names(roc.list) <- c('Original predictors (11)', 'Upregulated new predictors (67)')

ggroc(roc.list) + facet_wrap(~name) + theme(legend.position = 'none') + geom_text(data.labels, mapping = aes(0.5, 1, label= paste(label_AUC)), hjust = 1) + geom_text_repel(data.labels, mapping = aes(specificity, sensitivity, label = paste0(
  round(threshold, 1), ' (', 
  round(specificity, 1), '%, ', round(sensitivity, 1),'%)')
  ), nudge_y = -1, min.segment.length = 0, force = 3) + ggtitle('ROC curves for each GSEA predictor set') + theme_pubclean() + theme(legend.position = 'none')


Results_11$Prediction_GSEA <- ifelse(Results_11$nes > 1.6, 'transforming', 'untransforming')
# Results_114$Prediction_GSEA <- ifelse(Results_114$nes > 2, 'transforming', 'untransforming')
# Results_159$Prediction_GSEA <- ifelse(Results_159$nes > 1.9, 'transforming', 'untransforming')
# Results_99$Prediction_GSEA <- ifelse(Results_99$nes > 1.7, 'transforming', 'untransforming')
Results_67$Prediction_GSEA <- ifelse(Results_67$nes > 2, 'transforming', 'untransforming')


ggplot(Results_114[!Results_114$Protocol == 'standard (Clones)',], aes(x = Vector, y = nes, col = Design.x)) + ggbeeswarm::geom_quasirandom() + scale_color_manual(values = condition_colors) + facet_wrap(~Protocol, scales = 'free') + geom_hline(yintercept = 1.7, linetype = 'dashed') + theme_pubclean() + theme(legend.position = 'none', axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

ggplot(Results_11[!Results_11$Protocol == 'standard (Clones)',], aes(x = Vector, y = nes, col = Design.x)) + ggbeeswarm::geom_quasirandom() + scale_color_manual(values = condition_colors) + facet_wrap(~Protocol, scales = 'free') + geom_hline(yintercept = 1.6, linetype = 'dashed') + theme_pubclean() + theme(legend.position = 'none', axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + geom_hline(aes(yintercept = 1), col = 'red', linetype = 'longdash')

ggplot(Results_159[!Results_159$Protocol == 'standard (Clones)',], aes(x = Vector, y = nes, col = Design.x)) + ggbeeswarm::geom_quasirandom() + scale_color_manual(values = condition_colors) + facet_wrap(~Protocol, scales = 'free') + geom_hline(yintercept = 1.9, linetype = 'dashed') + theme_pubclean() + theme(legend.position = 'none', axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

ggplot(Results_67[!Results_67$Protocol == 'standard (Clones)',], aes(x = Vector, y = nes, col = Design.x)) + ggbeeswarm::geom_quasirandom() + scale_color_manual(values = condition_colors) + facet_wrap(~Protocol, scales = 'free') + geom_hline(yintercept = 2, linetype = 'dashed') + theme_pubclean() + theme(legend.position = 'none', axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))


# Use same ordering as SIF object
Results_11 <- Results_11[rownames(SIF[SIF$Sample_ID %in% Results_11$Sample_ID,]), ]
Results_114 <- Results_114[rownames(SIF[SIF$Sample_ID %in% Results_114$Sample_ID,]), ]
# Results_159 <- Results_159[rownames(SIF[SIF$Sample_ID %in% Results_159$Sample_ID,]), ]
Results_99 <- Results_99[rownames(SIF[SIF$Sample_ID %in% Results_99$Sample_ID,]), ]

# Export predictions
Results_merged <- SIF[SIF$Sample_ID %in% Results_11$Sample_ID, ]
Results_merged$GSEA_NES_11 <- Results_11$nes
Results_merged$GSEA_Prediction_11 <- Results_11$Prediction_GSEA
Results_merged$GSEA_NES_141 <- Results_114$nes
Results_merged$GSEA_Prediction_114 <- Results_114$Prediction_GSEA
# Results_merged$GSEA_NES_159 <- Results_159$nes
# Results_merged$GSEA_Prediction_159 <- Results_159$Prediction_GSEA
Results_merged$GSEA_NES_99 <- Results_99$nes
Results_merged$GSEA_Prediction_99 <- Results_99$Prediction_GSEA
Results_merged <- Results_merged[, -c(7, 8)]
write.csv(Results_merged, '/home/mmallee/GSEA_predictions_with99.csv')

wrong_pred_11 <- subset(Results_merged, subset= (!Class == GSEA_Prediction_11))
wrong_pred_99 <- subset(Results_merged, subset= (!Class == GSEA_Prediction_99))
wrong_pred_114 <- subset(Results_merged, subset= (!Class == GSEA_Prediction_114))

cm <- confusionMatrix(data = as.factor(Results_99$Class), as.factor(Results_99$Prediction_GSEA))
plt <- as.data.frame(cm$table)
plt$Prediction <- factor(plt$Prediction, levels=rev(levels(plt$Prediction)))
p <- ggplot(plt, aes(Prediction,Reference, fill= Freq)) +
geom_tile() + geom_text(aes(label=Freq)) +
scale_fill_gradient(low="white", high="#259999") +
labs(x = "Reference",y = "Prediction")+ ggtitle(paste('Confusion matrix 99'), subtitle = paste('Total samples: ', nrow(Results_99))) + theme_pubclean() + theme(legend.position = 'none')

p

plts <- lapply(c(11, 114, 99), function(geneset) {
  cm <- confusionMatrix(data = as.factor(Results_merged$Class), as.factor(Results_merged[[paste0('GSEA_Prediction_', geneset)]]))
  plt <- as.data.frame(cm$table)
  plt$Prediction <- factor(plt$Prediction, levels=rev(levels(plt$Prediction)))
  p <- ggplot(plt, aes(Prediction,Reference, fill= Freq)) +
  geom_tile() + geom_text(aes(label=Freq)) +
  scale_fill_gradient(low="white", high="#259999") +
  labs(x = "Reference",y = "Prediction")+ ggtitle(paste('Confusion matrix', geneset), subtitle = paste('Total samples: ', nrow(Results_merged[!is.na(Results_merged$Class), ]))) + theme_pubclean() + theme(legend.position = 'none')
  return(p)
})

wrap_plots(plts)


### investigate fc fluctuations
ok <- t(as.data.frame(map(FC_full, 1)))

ok <- apply(ok, 1, rank)

df3 <- merge(t(ok), SIF, by.x = 0, by.y = 'Sample_ID')
df3 <- reshape2::melt(df3, id = c(colnames(SIF)[-12], 'Row.names'))
ggplot(data = df3[df3$Design %in% c('A1_Mutagenic', 'A3_Safe') & df3$Protocol %in% c('published', 'standard', 'standard (JF)'), ], aes(x = variable, y = value, col = Design)) + ggbeeswarm::geom_quasirandom() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1), legend.position = 'top') + facet_wrap(~Design) + scale_color_manual(values = condition_colors)

```




### Report sampleinformation
```{r}
# Full report pData
## Only published data and new data from same protocol
pData_model <- pData_final[pData_final$Protocol %in% c('published'),]
## Remove test vectors
pData_model <- pData_model[-grep("test_vector", pData_model$Vector), ]

# Remove test/safe batches
safebatches <- unique(pData_model$Batch[pData_model$Design == 'A3_Safe'])
pData_model <- pData_model[!pData_model$Batch %in% safebatches,]

## Remove unknown class samples
pData_model <- pData_model[!pData_model$Design == 'A4_Unknown', ]
pData_model <- droplevels(pData_model)

write.csv(pData_model, '/home/mmallee/pData_train_412.csv', row.names = F) # Done
write.csv(pData_model, '/home/mmallee/pData_train_230.csv', row.names = F) # Done
write.csv(pData_model, '/home/mmallee/pData_258.csv', row.names = F) # Done
write.csv(pData_model, '/home/mmallee/pData_169.csv', row.names = F) # Done
write.csv(pData_model, '/home/mmallee/pData_154.csv', row.names = F) # Done

pData_412 <- read.csv('/home/mmallee/pData_train_412.csv')
pData_412$Batch <- as.factor(pData_412$Batch)
pData_230 <- read.csv('/home/mmallee/pData_train_230.csv')
pData_230$Batch <- as.factor(pData_230$Batch)
pData_258 <- read.csv('/home/mmallee/pData_258.csv')
pData_258$Batch <- as.factor(pData_258$Batch)
pData_169 <- read.csv('/home/mmallee/pData_169.csv')
pData_169$Batch <- as.factor(pData_169$Batch)
pData_154 <- read.csv('/home/mmallee/pData_154.csv')
pData_154$Batch <- as.factor(pData_154$Batch)

# SampleInformation file for testing original SAGA package
SampleInformation <- pData_model[pData_model$Protocol %in% c('standard', 'standard (JF)'), ]

SampleInformation$Group[SampleInformation$Design == 'A2_Mock'] <- 1
# Give each sample group nr 2:nsamples per batch
batches <- levels(as.factor(SampleInformation$Batch))
for (batch in batches) {
  nonmock <- SampleInformation[SampleInformation$Batch == batch & !SampleInformation$Design == 'A2_Mock',]
  SampleInformation[rownames(nonmock), "Group"] <- 2:(nrow(nonmock)+1)
}

SampleInformation <- SampleInformation[, c('Sample_ID', 'Filename', 'Group', 'Batch', 'Vector', 'Class')]
colnames(SampleInformation) <- c('SAMPLE_ID', 'Filename', 'Group', 'Batch', 'Vector', 'TrueLabel')
write.table(SampleInformation, 'SampleInformation.txt', sep = '\t', col.names = T, row.names = F)

SIF <- read.delim(paste(smplpath, "/SampleInformation.txt", 
    sep = ""), row.names = 1, header = TRUE, sep = "\t", 
    stringsAsFactors = F)

vector_spread <- as.data.frame.matrix(table(pData$Design, pData$Vector))
```




### Report figure 1 - GEX Analysis
```{r}
# 1a PCA of combined samples
raw_agilent <- read.table('/exports/immu-staallab/Micah/RAW_FullSagaSet700_quadruplicates_FINAL.txt', sep = '\t', header = T)[, -1]
eset_412 <- get_BC_eset(pData_412)
pca <- prcomp(t(eset_412))
pca_summary <- summary(pca)
pca_df <- data.frame(x = pca$x[, 1], y = pca$x[, 2]) %>% bind_cols(., pData_412)
p1a <- ggplot(data = pca_df, aes_string("x", "y", col = "Design")) +
  geom_point() +
  ggpubr::theme_pubclean() + theme(legend.key = element_rect(fill = NA)) + 
  ggtitle(label = paste0("PCA plot of combined samples")) + scale_color_manual(values = condition_colors) + xlab(label = paste0("PC1 - Explained Variance (", round(pca_summary$importance[2, 1] * 100, 1), "%)")) + ylab(label = paste0("PC2 - Explained Variance (", round(pca_summary$importance[2, 2] * 100, 1), "%)"))

# 1b, three panels

# Volcanoplots
topMvMoS_258 <- read.table('~/SAGA_GEX_Analysis/saga258/Toplist_FullSagaSet258_Mutagenic_vs_Mock_and_Safe_FINAL_batch.txt', header = T, sep = '\t')

topMvMoS_154 <- read.table('~/SAGA_GEX_Analysis/saga154/Toplist_FullSagaSet154_Mutagenic_vs_Mock_and_Safe_FINAL_batch.txt', header = T, sep = '\t')

topMvMoS_412 <- read.table('~/SAGA_GEX_Analysis/saga412/Toplist_FullSagaSet412_Mutagenic_vs_Mock_and_Safe_FINAL_batch.txt', header = T, sep = '\t')


top50_match <- topMvMoS_258[topMvMoS_258$X %in% topMvMoS_258$X[which(head(topMvMoS_258$X, 50) %in% head(topMvMoS_154$X, 50))], ]

hlgenes <- head(top50_match$GeneSymbol_FINAL, 15)

# Volcanoplot new data
p1b1 <- EnhancedVolcano(toptable = topMvMoS_258, x = 'logFC', y = 'adj.P.Val', lab = topMvMoS_258$GeneSymbol_FINAL, labSize = 4, max.overlaps = 10, widthConnectors = 0.75, drawConnectors = T, FCcutoff = 1.5, title = 'Differentially expressed probes', subtitle = 'Mutagenic vs Mock + Safe in unpublished data', caption = NULL, xlim = c(-5, 5), ylim = c(0, 40), selectLab = hlgenes, pCutoff = ) + theme_pubclean() + theme(legend.key = element_rect(fill = NA))

# Volcanoplot published data
p1b2 <- EnhancedVolcano(toptable = topMvMoS_154, x = 'logFC', y = 'adj.P.Val', lab = topMvMoS_154$GeneSymbol_FINAL, labSize = 4, max.overlaps = 10, widthConnectors = 0.75, drawConnectors = T, FCcutoff = 1.5, title = 'Differentially expressed probes', subtitle = 'Mutagenic vs Mock + Safe in published data', caption = NULL, xlim = c(-5, 5), ylim = c(0, 40), selectLab = hlgenes) + theme_pubclean() + theme(legend.key = element_rect(fill = NA))

p1b3 <- EnhancedVolcano(toptable = topMvMoS_412, x = 'logFC', y = 'adj.P.Val', lab = topMvMoS_412$GeneSymbol_FINAL, labSize = 4, max.overlaps = 10, widthConnectors = 0.75, drawConnectors = T, FCcutoff = 1, title = 'Differentially expressed probes', subtitle = 'Mutagenic vs Mock + Safe in combined data', caption = NULL, xlim = c(-5, 5), ylim = c(0, 40), selectLab = hlgenes) + theme_pubclean() + theme(legend.key = element_rect(fill = NA))


p1b <- (p1b1 + p1b2 + p1b3 + plot_layout(guides = 'collect')) & theme(legend.position = 'top')

# GSEA mock vs mutagenic + safe
library(fgsea)
eset.batch <- eset_412

sets <- readxl::read_excel(path = "~/Hannover_data/1-s2.0-S1525001621003233-mmc4.xlsx", sheet = 1, skip = 1, col_names = T)[, -2]
sets <- lapply(split(sets[, -1], sets[, 1]), unlist)
sets <- lapply(sets, function(x) {
  x[!is.na(x)]
})

mast_sets <- GSA.read.gmt(filename = '~/genesets.v2023.2.Hs_mast.gmt')
names(mast_sets$genesets) <- mast_sets$geneset.names
sets <- append(sets, mast_sets$genesets)
genes <- unique(unlist(sets))

# Make DataSet with one probe for each gene (select probe with highest variation)
eset <- data.frame(eset.batch, GeneSymbol = Annotation.known$GeneSymbol_FINAL)
eset.gsea <- subset(eset, toupper(eset$GeneSymbol) %in% genes) # take all probes with entries in GSEA-Sets
eset.sd <- apply(eset.gsea[1:length(eset.gsea) - 1], 1, sd)
eset.sd.order <- eset.gsea[order(eset.sd, decreasing = T), ]
eset.gsea <- subset(eset.sd.order, !duplicated(eset.sd.order$GeneSymbol))
matrix.gsea <- as.matrix(eset.gsea[, c(1:length(eset.gsea) - 1)])
Annotation.GSEA <- Annotation.known[row.names(eset.gsea), ]

# Filter for unique genes
genes <- genes[genes %in% toupper(Annotation.GSEA$GeneSymbol_FINAL)]

topMvMoS_412 <- topMvMoS_412[topMvMoS_412$X %in% rownames(Annotation.GSEA), ]

# Creating ranks
ranks <- topMvMoS_412$logFC[toupper(topMvMoS_412$GeneSymbol_FINAL) %in% genes]
names(ranks) <- toupper(topMvMoS_412$GeneSymbol_FINAL[toupper(topMvMoS_412$GeneSymbol_FINAL) %in% genes])

fgseaRes <- fgseaMultilevel(pathways = sets, ranks)

fgseaResTidy <- fgseaRes %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy <- fgseaResTidy[fgseaResTidy$padj < 0.05,]

p1c <- ggplot(fgseaResTidy, aes(reorder(pathway, NES), NES)) + geom_col(aes(fill = NES)) +
  coord_flip() + labs(x = "Pathway", y = "Normalized Enrichment Score", title = "GSEA results Mutagenic vs. Mock/Safe") + theme_pubclean() + scale_fill_gradient2(low = "#b2df8a", high = "#CB181D", midpoint = 0) + theme(legend.position = 'none')

(p1a + (p1b) + plot_layout(widths = c(1, 3))) / patchwork::free(p1c) + plot_layout(heights = c(1, 2)) + plot_annotation(tag_levels = 'A')


# Supplemental, foldchange vs foldchange
joined_fc <- inner_join(topMvMoS_154, topMvMoS_258, by = 'X')

joined_fc <- as.data.frame(joined_fc)
joined_fc$Sig <- "NS"
joined_fc$Sig[(joined_fc[["adj.P.Val.y"]] > 0.05 ) & (joined_fc[["adj.P.Val.x"]] < 0.05 )] <- "Significant in published"
joined_fc$Sig[(joined_fc[["adj.P.Val.y"]] < 0.05 ) & (joined_fc[["adj.P.Val.x"]] > 0.05 )] <- "Significant in unpublished"
joined_fc$Sig[(joined_fc[["adj.P.Val.y"]] < 0.05 ) & (joined_fc[["adj.P.Val.x"]] < 0.05 )] <- "Significant in both"

ggplot(joined_fc, mapping = aes(x = logFC.x, y = logFC.y, color = Sig)) + 
  geom_point(alpha = 0.5) +
  stat_cor(aes(x = logFC.x, y = logFC.y), label.x = 0) + 
  geom_smooth(method='lm', color = '#5b388a') + 
  geom_hline(yintercept = c(-1.5, 1.5), linetype = 'longdash', color = 'grey') +
  geom_vline(xintercept = c(-1.5, 1.5), linetype = 'longdash', color = 'grey') +
  labs(title = 'log2-Foldchanges in published and unpublished data', x = 'Log2 fold change in published data', y = 'Log2 fold change in unpublished data', subtitle = 'Mutagenic vs. Mock + Safe') + 
  theme_pubclean() + scale_color_identity(guide = 'legend') + scale_colour_manual(name = 'Sig', values = c(NS = 'grey30', 'Significant in unpublished' = 'forestgreen', 'Significant in published' = 'royalblue', 'Significant in both' = 'red2')) + guides(shape = 1) + 
  theme(legend.position = 'right')
```


### Report figure 2 - scRNA-seq
```{r}
### Main text:
# Figure 2: scRNA-seq
pData_258 <- read.csv('/home/mmallee/pData_258.csv')
real_seurat <- readRDS('/exports/immu-staallab/Micah/real_seurat.rds')
library(Seurat)
CIBERSORT_RESULTS <- read.table('/exports/immu-staallab/Micah/Deconvolution_412_final/CIBERSORTx_Results.txt', sep = '\t', header = T)
combined <- bind_cols(CIBERSORT_RESULTS, pData_412)
ct_colors <- c("#66a453", "#6c81d9", "#5b388a", "#8fb03d", "#c99f3b", "#c169ba", "#45c097", "#b74873", "#ab7434", "#bb4c41")
names(ct_colors) <- colnames(combined)[2:11]
names(ct_colors)[c(6, 7)] <- c('Mast_cell/Basophil', 'DC/Macrophage')
# A: UMAP of scRNA-seq data
p2a <- DimPlot(real_seurat, group.by = 'cell_type', cols = ct_colors, raster = F, label = T, shuffle = T, repel = T) + 
  theme_pubclean() + theme(legend.key = element_rect(fill = NA), legend.position = 'right') + ggtitle(label = 'Overview of cell types present in day 15 IVIM samples')

# Annotation reasoning?
annotation_features <- c('Fcer1a', 'Kit', 'Cma1', 'Tpsab1', 'Tpsb2', 'Car8', 'Hdc', 'Hgf', 'Perp', 'Cd38', 'Mmp19', 'Slc7a2', 'Mertk', 'Chil3', 'Stfa2', 'Stfa2l1', 'Chil1', 'Il1f9', 'Pf4', 'Gng11', 'Stmn1', 'Hist1h2ap', 'Top2a')

p2b <- DotPlot(real_seurat, features = annotation_features, dot.min = 0, col.max = 2, scale.min = 0, scale.max = 2, group.by = 'cell_type', col.min = 0) + ggtitle(label = 'Marker gene expressions used for annotation') + ggpubr::theme_cleveland() + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), plot.title = element_text(face = 'plain'), plot.tag = element_text(face = 'plain')) + scale_color_gradient(low = 'lightgrey', high = '#5b388a')


# Sample proportion for each cluster
p2c <- barplot.seurat(obj = real_seurat, id1 = 'Sample', 'cell_type') + xlab('Sample') + ylab('Percentage of cells') + ggtitle('Cell type composition per sample') + scale_fill_manual(values = ct_colors) + theme_pubclean()


# FACS differences
facs_results <- readxl::read_excel(path = '~/Master_Database.xlsx', sheet = 2, na = 'NA', range = readxl::cell_cols("A:F"))
facs_results_merged <-  plyr::join(pData_258, facs_results, by = 'Filename', type = 'inner')

colnames(facs_results_merged)[17:19] <- c('Myeloid', 'Lineage negative', 'Mast (FCER1a+)')

cell_types <- colnames(facs_results_merged)[17:19]
 differences <- facs_results_merged %>%
  group_by(Batch, Class) %>%
  summarise_at(cell_types, median) %>% 
  group_by(Batch) %>% 
  summarise(-across(cell_types, diff))
means_p_batch <- facs_results_merged %>% group_by(Batch) %>% summarise_at(cell_types, mean)
  means_p_batch <- means_p_batch[!means_p_batch$Batch == 58,]
ok <- melt(means_p_batch, id = c('Batch'))
p2d <- ggplot(data = melt(differences, id = c('Batch')), aes(x=variable, y = value, col = ok$value)) + ggbeeswarm::geom_quasirandom(size = 4) + xlab('Cell type') + ggtitle('Difference in FACS sort proportion per batch') + theme_pubclean() + guides(col = guide_colourbar(title = "Mean cell proportion")) + scale_color_gradient(low = "lightgrey", high = '#5b388a') + geom_hline(yintercept = 0, linetype = 'longdash', color = 'darkgrey') + theme(axis.line.y = element_line(arrow = grid::arrow(ends = "both"), colour = 'darkgrey')) + annotate("text", x = c(0.9, 0.9), y = c(-0.3, 0.3), label = c("Higher in untransforming", 'Higher in transforming')) + ylab('Difference in mean cell proportion') + ylim(c(-0.4, 0.4))

(p2a + p2b) / (p2c + p2d) + plot_annotation(tag_level = 'A')



# Supplemental, regress out cell cycle
real_seurat$CC.Difference <- real_seurat$S.Score - real_seurat$G2M.Score
real_seurat_cc <- ScaleData(real_seurat, vars.to.regress = "CC.Difference", features = rownames(real_seurat))

real_seurat_cc <- RunPCA(real_seurat_cc, npcs = 30, verbose = FALSE, features = rownames(real_seurat_cc))
ElbowPlot(real_seurat_cc)
real_seurat_cc <- RunUMAP(real_seurat_cc, reduction = "pca", dims = 1:10)

DimPlot(real_seurat_cc, group.by = 'cell_type', label = T) + ggtitle('Cell cycle regressed', 'Colored based on original clusters') + scale_color_manual(values = ct_colors) + theme_pubclean()
DimPlot(real_seurat, group.by = 'cell_type', label = T) + ggtitle('Cell cycle regressed') + scale_color_manual(values = ct_colors) + theme_pubclean()
```


### Report Figure 3 - Deconvolution results
```{r}
  # Overall deconvolution results
  # p3a
  names(ct_colors) <- colnames(combined)[2:11]
  

# FACS differences
pData_258 <- read.csv('/home/mmallee/pData_258.csv')
facs_results <- readxl::read_excel(path = '~/Master_Database.xlsx', sheet = 2, na = 'NA', range = readxl::cell_cols("A:F"))
facs_results_merged <-  plyr::join(pData_258, facs_results, by = 'Filename', type = 'inner')


  # differences in composition per batch
  # ?: GGbeeswarm of difference in cell type percentage per batch, gradient by absolute numbers
  cell_types <- colnames(combined)[2:11]
   differences <- combined %>%
    group_by(Batch, Class) %>%
    summarise_at(cell_types, median) %>% 
    group_by(Batch) %>% 
    summarise(-across(cell_types, diff))
  means_p_batch <- combined %>% group_by(Batch) %>% summarise_at(cell_types, mean)
    
  means_p_batch <- melt(means_p_batch, id = c('Batch'))
  p3a <- ggplot(data = melt(differences, id = c('Batch')), aes(x=variable, y = value, col = means_p_batch$value)) + 
    ggbeeswarm::geom_quasirandom() + 
    xlab('Cell type') + 
    ggtitle('Difference in estimated cell type proportion per batch') + 
    theme_pubclean() + 
    guides(col = guide_colourbar(title = "Mean cell proportion")) + 
    scale_color_gradient(low = "lightgrey", high = '#5b388a') + 
    geom_hline(yintercept = 0, linetype = 'longdash', color = 'darkgrey') + 
    theme(axis.line.y = element_line(arrow = grid::arrow(ends = "both"), colour = 'darkgrey'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), axis.ticks.y = element_blank()) + 
    annotate("text", x = c(3.9, 3.9), y = c(-0.35, 0.35), label = c("Higher in untransforming", 'Higher in transforming')) + 
    ylab('Difference in mean estimated cell proportion') + ylim(c(-0.4, 0.4))
  
  
  # Correlation with flow
  # facs_and_deconv <- merge(x = facs_results, y = combined_facs, by.x = 'Filename', by.y = 'Filename')
  facs_and_deconv <-  plyr::join(combined, facs_results, by = 'Filename', type = 'inner')
  colnames(facs_and_deconv)[31:33] <- c('Myeloid', 'Lineage_negative', 'FACS_Mast')
  facs_and_deconv$Deconvolution_Mast <- facs_and_deconv$Basophil + facs_and_deconv$Mast_cell + facs_and_deconv$Mast_cell.Basophil + facs_and_deconv$HSPC

p3b <- ggscatter(facs_and_deconv, x = "FACS_Mast", y = "Deconvolution_Mast", col= 'Batch', 
            add = "reg.line", add.params = list(color = "#5b388a", fill = "lightgray"), conf.int = TRUE) + 
  stat_cor(method = "pearson") + 
  ggtitle(label = 'Estimated mast cell composition vs. FACS mast cell composition') + 
  theme_pubclean() + 
  theme(legend.key = element_rect(fill = NA)) + 
  scale_color_manual(values = unname(ct_colors)[1:7]) + 
  labs(x = 'FACS Mast (Fcer1a+) proportion', y = 'Estimated mast cell proportion') + 
  lims(x = c(0, 1), y = c(0, 1)) + 
  coord_fixed(ratio = 1)

# SAGA prediction vs mast cell proportion
supplemental_facs_corr <- ggscatter(facs_and_deconv, x = "FACS_Mast", y = "Deconvolution_Mast", col= 'Batch', 
            add = "reg.line", add.params = list(color = "#5b388a", fill = "lightgray"), conf.int = TRUE) + 
  stat_cor(method = "pearson") + 
  ggtitle(label = 'Estimated mast cell composition vs. FACS mast cell composition') + 
  theme_pubclean() + 
  theme(legend.key = element_rect(fill = NA)) + 
  scale_color_manual(values = unname(ct_colors)[1:7]) + 
  labs(x = 'FACS Mast (Fcer1a+) proportion', y = 'Estimated mast cell proportion') + 
  facet_wrap(~Batch) + 
  lims(x = c(0, 1), y = c(0, 1))

# Print final figure
(p3a / p3b) + plot_annotation(tag_levels = 'A') + plot_layout(widths = 1)
```

### Report Figure 4 - Model testing
```{r}
# CV filtering
library(pROC)
ct_pseudo <- AggregateExpression(object = real_seurat, assays = 'RNA', return.seurat = T, group.by = 'cell_type')


df <- data.frame(sd = apply(ct_pseudo@assays$RNA$counts, 1, sd), mean = apply(ct_pseudo@assays$RNA$counts, 1, mean))
df$CV <- df$sd/df$mean
df$CV[df$mean == 0] <- 0

# First select genes based on total counts, to not have only genes with < 50 counts
counts <- ct_pseudo@assays$RNA$counts

df$counts <- apply(counts, 1, sum)

p4a <- ggplot(df, aes(x=log2(counts))) + geom_density() + geom_vline(xintercept = log2(50), color = '#CB181D') + ggtitle('Counts density \n(log2 transformed)') + ggplot(df, aes(x=CV)) + geom_density() + geom_vline(xintercept = 2, color = '#CB181D') + ggtitle('Coefficient of variance')


# Figure 5a RFE accuracies
rfe.73 <- readRDS('/home/mmallee/SAGA_Model_5.0_Full_dataset/rfe.FINAL.rds')

p4b <- ggplot(data = rfe.73[["results"]], aes(x = Variables, y = Accuracy)) + geom_line() + geom_point() + labs(title = 'RFE accuracy per feature subset', x = 'Number of variables retained', y = 'Accuracy (repeated cross validation)') + theme_pubclean() + geom_point(data = rfe.73[["results"]][48, ], aes(x = Variables, y = Accuracy), colour = "#CB181D", size = 2)


# GA accuracies
ga.73 <- readRDS('/home/mmallee/SAGA_Model_5.0_Full_dataset/gafinal.rds')

performance.external.FINAL <- ga.73$external
performance.external.FINAL <- arrange(performance.external.FINAL, Iter)
performance.FINAL <- ga.73$ga$internal
performance.FINAL$AccuracyExternal <- aggregate(performance.external.FINAL$Accuracy, by = list(Iter = performance.external.FINAL$Iter), mean)$x
performance.long.FINAL <- data.frame(Iter = c(performance.FINAL$Iter, performance.FINAL$Iter), Accuracy = c(performance.FINAL$AccuracyExternal, performance.FINAL$Accuracy), Group = c(rep("external", 40), rep("internal", 40)))


p4c <- ggplot(performance.long.FINAL, aes(Iter, Accuracy, col = Group)) +
  geom_point() +
  geom_smooth(span = 0.7, se = T) +
  theme_pubclean() +
  labs(title = 'Genetic algorithm internal and \nexternal accuracy', x = 'GA iteration', y = 'Accuracy')

# # Confusion matrix
Prediction_SVM.Caret <- read.csv('/home/mmallee/SAGA_Model_5.0_Full_dataset/SVM_test_73_predictions.csv')
Prediction_SVM.Caret$Protocol[Prediction_SVM.Caret$Protocol %in% c('standard', 'standard (JF)')] <- 'Unpublished'
Prediction_SVM.Caret$Protocol[Prediction_SVM.Caret$Protocol == 'published'] <- 'Published'
cm <- confusionMatrix(data = as.factor(Prediction_SVM.Caret$Class), as.factor(Prediction_SVM.Caret$Prediction.SVM.Caret))
plt <- as.data.frame(cm$table)
plt$Prediction <- factor(plt$Prediction, levels=rev(levels(plt$Prediction)))

p4d1 <- ggplot(plt, aes(Prediction,Reference, fill= Freq)) +
  geom_tile() +
  geom_text(aes(label=Freq)) +
  scale_fill_gradient(low="white", high="#b2df8a") +
  labs(x = "Reference",y = "Prediction") +
  ggtitle('Confusion matrix', subtitle = paste('Total samples: ', nrow(Prediction_SVM.Caret))) +
  theme_pubclean() +
  theme(legend.position = 'none')

# ROC 
roc.73 <- roc(Prediction_SVM.Caret$Class, Prediction_SVM.Caret$transforming, percent = TRUE, 
              levels=c("untransforming","transforming"), plot = T, auc.polygon = F, 
              max.auc.polygon = F, col = "#686868", lwd = 4, grid = F, add = F,
              print.auc = T, print.thres = "best")
data.labels <- coords(roc.73)

roc.73$auc

data.labels <- coords(roc.73, 'best', input = 'threshold', ret = c('threshold', 'specificity', 'sensitivity'))
data.labels <- rbind(data.labels, coords(roc.73, 0.65, input = 'threshold', ret = c('threshold', 'specificity', 'sensitivity')))

data.labels <- coords(roc.73, 0.65, input = 'threshold', ret = c('threshold', 'specificity', 'sensitivity'))


p4d <- ggroc(roc.73) + 
  theme(legend.position = 'none') + 
  geom_text(data.labels, mapping = aes(0.5, 1, label = paste('AUC =', round(roc.73$auc, 2))), hjust = 1) + 
  geom_text_repel(data.labels, mapping = aes(specificity, sensitivity, label = paste0(round(threshold, 2), ' (', round(specificity, 1), '%, ', round(sensitivity, 1),'%)')), min.segment.length = 0, position = position_nudge_repel(x = 30, y = -15)) + 
  labs(title = 'ROC curve for SAGA-SVM \ntest set prediction', x = 'Specificity', y = 'Sensitivity') + theme_pubclean() + theme(legend.position = 'none')

# Beeswarm of test set predictions
p4e <- ggplot(Prediction_SVM.Caret, mapping = aes(x = Vector, y = transforming, color = Design)) + 
  ggbeeswarm::geom_quasirandom() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + 
  facet_wrap(~Protocol, scales = 'free') + 
  ylim(c(0, 1)) + 
  scale_color_manual(values = condition_colors) + 
  theme_pubclean() + 
  theme(legend.key = element_rect(fill = NA), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + 
  labs(title = 'Test set predictions', x = 'Vector type', y = 'Transforming prediction') + 
  geom_hline(yintercept = 0.65, alpha = 0.5, linetype = 'dashed', color = '#CB181D')

# Wrap plots
p4a / (p4b | p4c | p4d) / patchwork::free(p4e) + plot_layout(heights = c(1, 1,2)) + plot_annotation(tag_levels = 'A')
```

### Report Figure 5 - Final model performance estimation
```{r}
# Also include genes
rfe.final <- readRDS('/home/mmallee/SAGA_Model_5.0_Full_dataset/Final_412/rfe.FINAL.rds')
gafinal <- readRDS('/home/mmallee/SAGA_Model_5.0_Full_dataset/Final_412/gafinal.rds')


# 6A RFE accuracies
p6a <- ggplot(data = rfe.final[["results"]], aes(x = Variables, y = Accuracy)) + geom_line() + geom_point() + labs(title = 'RFE accuracy per feature subset', x = 'Number of variables retained', y = 'Accuracy (repeated cross validation)') + theme_pubclean() + geom_point(data = rfe.final[["results"]][48, ], aes(x = Variables, y = Accuracy), colour = "#CB181D", size = 2)


# GA accuracies
performance.external.FINAL <- gafinal$external
performance.external.FINAL <- arrange(performance.external.FINAL, Iter)
performance.FINAL <- gafinal$ga$internal
performance.FINAL$AccuracyExternal <- aggregate(performance.external.FINAL$Accuracy, by=list(Iter=performance.external.FINAL$Iter),mean)$x
performance.long.FINAL <- data.frame(Iter = c(performance.FINAL$Iter,performance.FINAL$Iter), Accuracy = c(performance.FINAL$AccuracyExternal,performance.FINAL$Accuracy), Group=c(rep("external",40),rep("internal",40)))


p6b <- ggplot(performance.long.FINAL, aes(Iter, Accuracy, col = Group)) +
  geom_point() +
  geom_smooth(span = 0.7, se = T) +
  theme_pubclean() +
  labs(title = 'Genetic algorithm internal and external accuracy', x = 'GA iteration', y = 'Accuracy')

real_seurat <- readRDS('/exports/immu-staallab/Micah/real_seurat.rds')
predictors_ann <- Annotation.known[gafinal$optVariables, ]
library(Seurat)
ct_colors <- c("#66a453", "#6c81d9", "#5b388a", "#8fb03d", "#c99f3b", "#c169ba", "#45c097", "#b74873", "#ab7434", "#bb4c41")
CIBERSORT_RESULTS <- read.table('/exports/immu-staallab/Micah/Deconvolution_412_final/CIBERSORTx_Results.txt', sep = '\t', header = T)
combined <- bind_cols(CIBERSORT_RESULTS, pData_412)
names(ct_colors) <- colnames(combined)[2:11]
names(ct_colors)[c(6, 7)] <- c('Mast_cell/Basophil', 'DC/Macrophage')
real_seurat <- ScaleData(real_seurat, do.scale = T, do.center = T)
p6c <- DoHeatmap(object = real_seurat, features = predictors_ann$GeneSymbol_FINAL, slot = 'scale.data', group.colors = ct_colors, angle = 90, label = F) + labs(title = 'Expression of predictors across scRNA-seq clusters', disp.min = -1, disp.max = 1)

DotPlot(real_seurat, features = predictors_ann$GeneSymbol_FINAL, group.by = 'cell_type', assay = 'RNA')

(p6a | p6b) / p6c + plot_layout(heights = c(1, 5)) + plot_annotation(tag_levels = 'A')

# Write final predictors to text file
f <- file("Predictors_74.txt")
writeLines(predictors_ann$GeneSymbol_FINAL, f)
close(f)
```


### Report Figure 6 - SAGA-GSEA
```{r}
# PCA of unbatchcorrected data
pData_firstthree <- pData_412[pData_412$Batch %in% c(1, 2, 53), ]
# Format as EListRaw
E <- raw_agilent[, pData_firstthree$Sample_ID]
E <- new("EListRaw", list(E = E, genes = data.frame(ProbeName = raw_agilent$ProbeID), targets = colnames(E), source = 'agilent.median'))
eset.rma <- normalizeBetweenArrays(object = E, method = "quantile")
eset.rma <- avereps(x = eset.rma, ID = E$genes$ProbeName) 
eset.rma_m <- eset.rma$E
eset.rma_m <- eset.rma_m[rownames(Annotation.known), ]

library(ggforce)

pca <- prcomp(t(eset.rma_m))
pca_summary <- summary(pca)
pca_df <- data.frame(x = pca$x[, 1], y = pca$x[, 2]) %>% bind_cols(., pData_firstthree)
p6a <- ggplot(data = pca_df, aes_string("x", "y", col = "Design")) +
  geom_point() +
  ggpubr::theme_pubclean() + theme(legend.key = element_rect(fill = NA)) + 
  labs(title = 'PCA plot of three batches', subtitle = 'Without batch correction', x = paste0("PC1 - Explained Variance (", round(pca_summary$importance[2, 1] * 100, 1), "%)"), y = paste0("PC2 - Explained Variance (", round(pca_summary$importance[2, 2] * 100, 1), "%)")) + 
  scale_color_manual(values = condition_colors) + 
  geom_mark_ellipse(aes(group = Batch, col = Batch, label = paste0('Batch: ', Batch)))






# GSEA plot of predictors
p6b1 <- plotEnrichment(pathway = SAGA.CORE$top67, stats = stats <- SAGA.GSEA$simulations$Group.2.fc$fc.hr) + labs(title = 'GSEA plot new predictors - Mutagenic vs Mock', subtitle = 'sample 7652', x = 'Rank', y = 'Enrichment Score (ES)')
p6b2 <- plotEnrichment(pathway = SAGA.CORE$top67, stats = SAGA.GSEA$simulations$Group.4.fc$fc.hr) + labs(title = 'GSEA plot new predictors - Mock vs Safe', subtitle = 'sample 7654', x = 'Rank', y = 'Enrichment Score (ES)')
p6b <- p6b1 / p6b2

# ROC of predictions
SIF <- read.csv('/home/mmallee/pData_train_412.csv', header = T)
SIF$Protocol[SIF$Protocol %in% c('standard', 'standard (JF)')] <- 'Unpublished'
SIF$Protocol[SIF$Protocol == 'published'] <- 'Published'
batches <- levels(as.factor(SIF$Batch)) # how many assays / batches 
batches <- batches[-which(batches == 11)]

results_true <- lapply(batches, function(batch){
  read.table(file = paste("~/SAGA_GSEA_results/Results_SAGA.GSEA_Batch_", batch,".txt", sep = ""), sep = "\t", header = T)
})
results_all <- data.table::rbindlist(results_true)

results_top11 <- results_all[results_all$geneSet == 'top11',]
results_top67 <- results_all[results_all$geneSet == 'top67',]

Results_11 <- merge(SIF, results_top11, by.x = 0, by.y = "SampleID", all.x = F)
Results_67 <- merge(SIF, results_top67, by.x = 0, by.y = "SampleID", all.x = F)
rownames(Results_11) <- Results_11$Row.names
rownames(Results_67) <- Results_67$Row.names


rm(results_top11, results_top67, results_true, results_all)

roc.11 <- roc(Results_11$Class,                    
                  Results_11$nes,             
                  percent = TRUE, levels=c("untransforming","transforming"),
                  plot = F, auc.polygon = F, max.auc.polygon = F, col = "#686868", lwd = 4, grid = F, add = F,
                  print.auc = T, print.thres = "best")

roc.67 <- roc(Results_67$Class,                    
                  Results_67$nes,             
                  percent = TRUE, levels=c("untransforming","transforming"),
                  plot = F, auc.polygon = F, max.auc.polygon = F, col = "#686868", lwd = 4, grid = F, add = F,
                  print.auc = T, print.thres = "best")

roc.list <- list(roc.11, roc.67)

# extract auc
roc.list %>% 
  map(~tibble(AUC = .x$auc)) %>% data.table::rbindlist() -> data.auc
data.auc$name <- c('Original predictors (11)', 'Upregulated new predictors (67)')

# generate labels labels
data.labels <- cbind(data.auc, lapply(roc.list, coords, 'best', input = 'threshold', ret = c("threshold", 'specificity', 'sensitivity')) %>% data.table::rbindlist())

names(roc.list) <- c('Original predictors (11)', 'Upregulated new predictors (67)')

p6c <- ggroc(roc.list) + 
  geom_text_repel(data.labels, mapping = aes(c(50, 25), y = 20, label = paste('AUC = ', round(AUC, 2))), hjust = 1) + 
  geom_text_repel(data.labels, mapping = aes(specificity, sensitivity, 
                                             label = paste0(round(threshold, 1), ' (', 
                                                            round(specificity, 1), '%, ', 
                                                            round(sensitivity, 1),'%)')), 
                  nudge_y = c(15, -5), min.segment.length = 0, force = 3, nudge_x = c(5, 15)) + 
  ggtitle('ROC curves for the original and new GSEA predictor set') + 
  theme_pubclean() + 
  geom_abline(intercept = c(100, 100), color = 'darkgray', linetype = 'dashed') +
  scale_color_manual(values = c('Original predictors (11)' = '#bb4c41', 'Upregulated new predictors (67)' = '#5b388a'))

# beeswarm of predictions
p6d <- ggplot(Results_67[!Results_67$Protocol == 'standard (Clones)',], aes(x = Vector, y = nes, col = Design.x)) +
  ggbeeswarm::geom_quasirandom() + 
  scale_color_manual(values = condition_colors) + 
  facet_wrap(~Protocol, scales = 'free_x') +
  geom_hline(yintercept = 2, linetype = 'dashed') + 
  theme_pubclean() + 
  theme(legend.position = 'none', axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + 
  labs(title = 'SAGA-GSEA predictions', subtitle = 'Using new predictor set', y = 'Normalized Enrichment Score (NES)')

# Print full figure:
(p6a | p6b) / (p6c | p6d) + plot_annotation(tag_levels = 'A')

```


### Create test/train set
```{r}
test_batches <- c(3, 4, 16, 17, 32, 41, 47, 57, 59, 66, 71, 73)
pData_test_73 <- pData_412[pData_412$Batch %in% test_batches,]
table(pData_test_73$Design)
# 39 mutagenic, 28 mock, 6 safe
pData_train_339 <- pData_412[!pData_412$Batch %in% test_batches,]

write.csv(pData_test_73, '/home/mmallee/pData_test_73.csv', row.names = F) # Done
write.csv(pData_train_339, '/home/mmallee/pData_train_339.csv', row.names = F) # Done

eset_339 <- get_BC_eset(droplevels(pData_train_339))
# Write data
write.table(eset_339, file = "/exports/immu-staallab/Micah/ESET_RMA_COMBAT_FullSagaSet339_FINAL.txt", sep = "\t", row.names = TRUE, col.names = NA)
```





### Cell surface predictors
```{r}
surface_protein_atlas <- readxl::read_excel("CPA.xlsx", sheet = "Table B")
gafinal <- readRDS('/home/mmallee/SAGA_Model_5.0_Full_dataset/Final_412/gafinal.rds')

predictors_ann <- Annotation.known[gafinal$optVariables, ]
hits <- merge(y = surface_protein_atlas, x = predictors_ann, by.y = 'ENTREZ gene symbol', by.x = 'GeneSymbol_FINAL')

predictors_ann <- cbind(ProbeID = rownames(predictors_ann), predictors_ann)

writexl::write_xlsx(list(predictors_ann, hits), 'Supplemental_table_S7_Final_74_predictors.xlsx')
```


### Report figures
```{r}
# Piechart samples
data <- pData_412
data <- data %>% 
  group_by(Design) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(Design))
data$label <- scales::percent(data$per)

ggplot(data=data)+
  geom_bar(aes(x = "", y = per, fill = Design), stat="identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  geom_text(aes(x = 1, y = cumsum(per) - per/2, label = label)) + 
  scale_fill_manual(values = condition_colors)


ggplot(pData_412, aes(x="", fill = Design, y = Design)) + geom_bar(stat = 'identity', width = 1) + coord_polar('y', start = 0) + geom_text(aes(label = Design))


df <- cbind(ProbeID = rownames(Annotation.known), Annotation.known)

writexl::write_xlsx(df, 'Supplemental_Table_S1_Probe_Annotation.xlsx')



pData_412
pData_test_73 <- read.csv('/home/mmallee/pData_test_73.csv')
pData_train_339 <- read.csv('/home/mmallee/pData_train_339.csv')
writexl::write_xlsx(list(pData_412, pData_test_73, pData_train_339), 'Supplemental_Table_S4_Test_train_split.xlsx')


unknown_markers <- FindMarkers(real_seurat, ident.1 = 'Unknown', only.pos = T, min.diff.pct = 0.3)
unknown_markers <- cbind(Gene = rownames(unknown_markers), unknown_markers)

writexl::write_xlsx(unknown_markers, 'Supplemental_Table_S5_Unknown_DEG.xlsx')

gafinal <- readRDS('/home/mmallee/SAGA_Model_5.0_Full_dataset/Final_412/gafinal.rds')
predictors_ann <- Annotation.known[gafinal$optVariables, ]
```

